---
title: "Pre-processing Taxonomic MetaPhlAn4 All samples - Yogurt study"
output: 
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
---


```{r}
library(phyloseq)
library(dplyr)
library(readxl)
library(tidyr)
library(ggplot2)
library(mia)
library(microbiome)
library(patchwork)
library(ggExtra)

dir <- "GitHub_repository/IPE_yogurt_rolledOat_microbiome/"
dir1 <- paste(dir, "DataPreparation/Microbiome/", sep="")

source(file = paste(dir, "R functions/Utils.R", sep=""))
source(file = paste(dir, "R functions/Utils_metagenomics.R", sep=""))
source(file = paste(dir, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir, "R functions/Utils_crossOver.R", sep=""))
```

```{r}
cols_Inter_treat <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")
```


# Loading tables
## Metadata
```{r}
metadata <- read_excel(paste(dir1, "Metadata.xlsx", sep=""), sheet=1)
metadata <- as.data.frame(metadata)

metadata$Sample_ID <- gsub("-", "", metadata$Sample_ID)
metadata$Treatment <- case_when(metadata$Timepoint %in% c("T1", "T3") ~ "Before",
                                metadata$Timepoint %in% c("T2", "T4") ~ "After",
                                TRUE ~ NA_character_)
metadata$Intervention <- case_when(metadata$Timepoint %in% c("T1", "T2") & metadata$Gruppe == "A" ~ "Yogurt",
                                   metadata$Timepoint %in% c("T3", "T4") & metadata$Gruppe == "A" ~ "YogurtOatmeal",
                                   metadata$Timepoint %in% c("T1", "T2") & metadata$Gruppe == "B" ~ "YogurtOatmeal",
                                   metadata$Timepoint %in% c("T3", "T4") & metadata$Gruppe == "B" ~ "Yogurt",
                                TRUE ~ NA_character_)

metadata$Inter_treat <- paste(metadata$Treatment, metadata$Intervention, sep=".")
metadata$Inter_treat <- factor(metadata$Inter_treat, levels = c("Before.Yogurt", "After.Yogurt", "Before.YogurtOatmeal", "After.YogurtOatmeal"), ordered=T)
metadata$Treatment <- factor(metadata$Treatment, levels = c("Before", "After"), ordered=T)
metadata$Timepoint <- factor(metadata$Timepoint, levels = c("T1", "T2", "T3", "T4"), ordered=T)
metadata$Gruppe <- as.factor(metadata$Gruppe)
metadata$Participant_ID <- as.factor(metadata$Participant_ID)
metadata$Geschlecht <- as.factor(metadata$Geschlecht)
metadata$Intervention <- as.factor(metadata$Intervention)

rownames(metadata) <- metadata[, "Sample_ID"]
```

```{r}
dim(metadata) # 440   11
```


## Species table

```{r}
table_rel_sp <- read.table("exp86-freiburg/merged_abundance_table_ALL_species.txt", header=T, row.names = "clade_name")
table_rel_sp_TAX <- read.table("exp86-freiburg/merged_abundance_table_ALL_species_TAX.txt", header=T, row.names = "clade_name")
dim(table_rel_sp) #  1378  440
summary(colSums(table_rel_sp))
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   #  100     100     100     100     100     100 
```


```{r}
table_rel_SGB <- read.table("exp86-freiburg/merged_abundance_table_ALL_SGB.txt", header=T, row.names = "clade_name")
table_rel_SGB_TAX <- read.table("exp86-freiburg/merged_abundance_table_ALL_SGB_TAX.txt", header=T, row.names = "clade_name")
dim(table_rel_SGB) #  1875  440
summary(colSums(table_rel_sp))
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   #  100     100     100     100     100     100 

# Change rownames to fit with the phylogenetic tree
#rownames(table_rel_SGB) <- gsub('SGB', '', rownames(table_rel_SGB))
rownames(table_rel_SGB) <- gsub('_group', '', rownames(table_rel_SGB))
```


## Taxonomic description
```{r}
taxa_table_sp <- data.frame(rownames(table_rel_sp_TAX))
taxa_table_sp <- separate(taxa_table_sp, col=1, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep="\\|")
taxa_table_sp <- gsub('.__', '', as.matrix(taxa_table_sp))
rownames(taxa_table_sp) <- taxa_table_sp[,"Species"]
```

```{r}
taxa_table_SGB <- data.frame(rownames(table_rel_SGB_TAX))
taxa_table_SGB <- separate(taxa_table_SGB, col=1, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "SGB"), sep="\\|")
taxa_table_SGB <- gsub('.__', '', as.matrix(taxa_table_SGB))
rownames(taxa_table_SGB) <- taxa_table_SGB[,"SGB"]

# Change rownames to fit with the phylogenetic tree
# rownames(taxa_table_SGB) <- gsub('SGB', '', rownames(taxa_table_SGB))
rownames(taxa_table_SGB) <- gsub('_group', '', rownames(taxa_table_SGB))
```


## Phylogenetic tree
```{r paged.print=FALSE}
tree <- read_tree("mpa_vJan21_CHOCOPhlAnSGB_202103.nwk.txt")
tree$tip.label <- paste("SGB", tree$tip.label, sep="")
as_tibble(tree)
```


# Creating Phyloseq objects

## physeq_rel_sp

```{r}
OTU = otu_table(table_rel_sp, taxa_are_rows = TRUE)
TAX = tax_table(taxa_table_sp)
SAM = sample_data(metadata)
dim(OTU)# [1] 1378  440
```

```{r}
inters_names <- intersect(colnames(OTU), rownames(metadata))
cat("Samples present in Metaphlan and absent from Metadata: ")
colnames(OTU)[!(colnames(OTU) %in% inters_names)]

cat("Samples present in Metadata and absent from Metaphlan: ")
rownames(metadata)[!(rownames(metadata) %in% inters_names)]
```

```{r}
physeq_rel_sp = phyloseq(OTU, TAX, SAM)
physeq_rel_sp 
```


```{r}
# Print sample / taxa lost when creating phyloseq object
if(!ncol(OTU) == ncol(otu_table(physeq_rel_sp))){
  inters_names <- intersect(colnames(OTU), colnames(otu_table(physeq_rel_sp)))
  cat("Phyloseq sample lost: ")
  colnames(OTU)[!(colnames(OTU) %in% inters_names)] %>% print()
}
if(!nrow(OTU) == nrow(otu_table(physeq_rel_sp))){
  inters_names <- intersect(rownames(OTU), rownames(otu_table(physeq_rel_sp)))
  cat("Phyloseq taxa lost: ")
  rownames(OTU)[!(rownames(OTU) %in% inters_names)] %>% print()
}
```

## physeq_rel_SGB
```{r message=FALSE}
OTU = otu_table(table_rel_SGB, taxa_are_rows = TRUE)
TAX = tax_table(taxa_table_SGB)
SAM = sample_data(metadata)
dim(OTU) # 1875  440
```

```{r}
inters_names <- intersect(colnames(OTU), rownames(metadata))
cat("Samples present in Metaphlan and absent from Metadata: ")
colnames(OTU)[!(colnames(OTU) %in% inters_names)]

cat("Samples present in Metadata and absent from Metaphlan: ")
rownames(metadata)[!(rownames(metadata) %in% inters_names)]
```


```{r message=FALSE}
physeq_rel_SGB = phyloseq(OTU, TAX, SAM, tree)
physeq_rel_SGB
```

```{r}
# Print sample / taxa lost when creating phyloseq object
if(!ncol(OTU) == ncol(otu_table(physeq_rel_SGB))){
  inters_names <- intersect(colnames(OTU), colnames(otu_table(physeq_rel_SGB)))
  cat("Phyloseq sample lost: ")
  colnames(OTU)[!(colnames(OTU) %in% inters_names)] %>% print()
}
if(!nrow(OTU) == nrow(otu_table(physeq_rel_SGB))){
  inters_names <- intersect(rownames(OTU), rownames(otu_table(physeq_rel_SGB)))
  cat("Phyloseq taxa lost: ")
  rownames(OTU)[!(rownames(OTU) %in% inters_names)] %>% print()
}
# Phyloseq taxa lost: [1] "EUK4932"   "EUK944036"
```


# Filtering

## physeq_rel_sp
```{r}
table(tax_table(physeq_rel_sp)[, "Kingdom"])
  # Archaea  Bacteria Eukaryota 
  #       9      1367         2 
```

### Description of the Eukaryota and Archea

```{r paged.print=FALSE}
# Eukaryota 
euk <- subset(as.data.frame(tax_table(physeq_rel_sp)), Kingdom=="Eukaryota")$Species
euk
# [1] "Saccharomyces_cerevisiae"  "Blastocystis_sp_subtype_1"
```


```{r paged.print=FALSE}
# Archaea  
arch <- subset(as.data.frame(tax_table(physeq_rel_sp)), Kingdom=="Archaea")$Species
arch
# [1] "Methanobrevibacter_smithii"                    "Methanosphaera_stadtmanae"                    
# [3] "Methanomassiliicoccaceae_unclassified_SGB380"  "GGB781_SGB1024"                               
# [5] "Candidatus_Methanomethylophilus_alvus"         "Candidatus_Methanomassiliicoccus_intestinalis"
# [7] "Methanomassiliicoccales_archaeon"              "GGB1567_SGB2154"                              
# [9] "GGB260_SGB362"      
```


```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
taxa <- c(euk, arch)

pp_list <- list()
absum <- abmax <- c()
DF <- as.data.frame(sample_data(physeq_rel_sp))
for(i in seq(length(taxa))){
  
  DF$abundance <- get_sample(physeq_rel_sp, taxa[i])
  absum <- c(absum, sum(DF$abundance > 0))
  abmax <- c(abmax, max(DF$abundance))
  
  pp_list[[i]] <- plot_sample_measure_CO(DF, "abundance", taxa[i],  x="Timepoint", colorInter = T, trans = "psd_log", print_table = F)  
  

}

wrap_plots(pp_list, nrow=3) 
```

```{r paged.print=FALSE}
data.frame(number_samples=absum, maximal_abundance=abmax, row.names = taxa)
#                                               number_samples maximal_abundance
# Saccharomyces_cerevisiae                                   7           0.00436
# Blastocystis_sp_subtype_1                                 19           0.40002
# Methanobrevibacter_smithii                               100           2.04360
# Methanosphaera_stadtmanae                                 25           0.50993
# Methanomassiliicoccaceae_unclassified_SGB380               1           0.05871
# GGB781_SGB1024                                            17          11.13196
# Candidatus_Methanomethylophilus_alvus                     11           0.47250
# Candidatus_Methanomassiliicoccus_intestinalis              3           0.18422
# Methanomassiliicoccales_archaeon                           4           0.37863
# GGB1567_SGB2154                                            4           0.36372
# GGB260_SGB362                                              3           0.53408
```

### Prevalence and abundance
```{r}
rel_ab <- abundances(physeq_rel_sp, transform = "compositional")*100 # Retrieves the taxa x samples table from phyloseq object after additional transformation (here normalisation)
total_abundance <- rowSums(rel_ab) 
max_abundance <- apply(rel_ab, 1, max)
preval <- prevalence(physeq_rel_sp)

thresh_abond=0.01
thresh_prev= 0.03 
df <- data.frame(Species= names(total_abundance), total_abundance=total_abundance, max_abundance=max_abundance, preval=preval)
```

```{r}
pp <- ggplot(df, aes(max_abundance, preval)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  geom_vline(xintercept = thresh_abond, color = "red", linetype = "dotted") +
  geom_hline(yintercept = thresh_prev, color = "red", linetype = "dotted")  + 
   theme_bw()
pp1 <- ggMarginal(pp, type="density") 

pp2 <- pp +
  scale_x_log10(labels = scales::label_number(), name = "Maximum Abundance") +
  scale_y_log10(
   labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
   name = "Prevalence (%)")
pp2 <- ggMarginal(pp2, type="density") 
```

```{r fig.height=7, fig.width=12}
wrap_plots(pp1, pp2)
```



```{r}
# Function in Utils_metagenomics.R
cleanUp1 = phyloseq_CleanUp(physeq_rel_sp, thresh_abond = 0, thresh_prev =0) # Cleaning out non bacterial taxa
cleanUp2 = phyloseq_CleanUp(physeq_rel_sp, thresh_abond = thresh_abond, thresh_prev = thresh_prev) # Cleaning out non bacterial and low abundant or prevalent taxa 
```

```{r paged.print=FALSE}
cleanUp1$df_filt
#         dim_data del_noSample del_abond<0% del_prev<0% del_taxaType dim_dataClean
# OTU         1378            0            0           0           11          1367
# Samples      440            0            0           0            0           440
```

```{r paged.print=FALSE}
cleanUp2$df_filt # 1377-662-17 = 698
#         dim_data del_noSample del_abond<0.01% del_prev<0.03% del_taxaType dim_dataClean
# OTU         1378            0             214            662           11           699
# Samples      440            0               0              0            0           440
```

```{r}
# Details of the deleted taxa
deleted_taxa_sp <- as.vector(tax_table(cleanUp2$deleted_taxa)[, "Species"])
table(tax_table(cleanUp2$deleted_taxa)[, "Kingdom"])
  # Archaea  Bacteria Eukaryota 
  #       9       668         2 
```


```{r}
physeq_rel_sp_Clean1 = prune_taxa(taxa_sums(cleanUp1$clean_subset) > 0,  cleanUp1$clean_subset)
physeq_rel_sp_Clean2 = prune_taxa(taxa_sums(cleanUp2$clean_subset) > 0,  cleanUp2$clean_subset)
#physeq_rel_sp_CleanR = transform_sample_counts(physeq_rel_sp_Clean, function(x) 100 * x/sum(x))
```


```{r}
rel_ab <- abundances(physeq_rel_sp_Clean2, transform = "identity")
max_abundance <- apply(rel_ab, 1, max)
preval <- prevalence(physeq_rel_sp_Clean2)
df <- data.frame(max_abundance=max_abundance, preval=preval)

pp <- ggplot(df, aes(max_abundance, preval)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  geom_vline(xintercept = thresh_abond, color = "red", linetype = "dotted") +
  geom_hline(yintercept = thresh_prev, color = "red", linetype = "dotted")  + 
   theme_bw() +
  scale_x_log10(labels = scales::label_number(), name = "Maximum Abundance") +
  scale_y_log10(
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)") 

pp <- ggMarginal(pp, type="density") 
```

```{r}
pp
```

## physeq_rel_SGB
```{r}
table(tax_table(physeq_rel_SGB)[, "Kingdom"])
```


```{r}
rel_ab <- abundances(physeq_rel_SGB, transform = "compositional")*100
total_abundance <- rowSums(rel_ab) 
max_abundance <- apply(rel_ab, 1, max)
preval <- prevalence(physeq_rel_SGB)

thresh_abond=0.01
thresh_prev= 0.03
df <- data.frame(Species= names(total_abundance), total_abundance=total_abundance, max_abundance=max_abundance, preval=preval)

pp <- ggplot(df, aes(max_abundance, preval)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  geom_vline(xintercept = thresh_abond, color = "red", linetype = "dotted") +
  geom_hline(yintercept = thresh_prev, color = "red", linetype = "dotted")  + 
   theme_bw()
pp1 <- ggMarginal(pp, type="density") 

pp2 <- pp +
  scale_x_log10(labels = scales::label_number(), name = "Maximum Abundance") +
  scale_y_log10(
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)") 
pp2 <- ggMarginal(pp2, type="density") 
```


```{r fig.height=7, fig.width=12}
wrap_plots(pp1, pp2)
```


```{r message=FALSE}
# Function in Utils_metagenomics.R
cleanUp1 = phyloseq_CleanUp(physeq_rel_SGB, thresh_abond = 0, thresh_prev =0) # Cleaning out non bacterial taxa
cleanUp2 = phyloseq_CleanUp(physeq_rel_SGB, thresh_abond = thresh_abond, thresh_prev = thresh_prev) # Cleaning out non bacterial and low abundant taxa (no filtering on prevalence)
```

```{r paged.print=FALSE}
cleanUp1$df_filt
#         dim_data del_noSample del_abond<0% del_prev<0% del_taxaType dim_dataClean
# OTU         1873            0            0           0            9          1864
# Samples      440            0            0           0            0           440
```

```{r paged.print=FALSE}
cleanUp2$df_filt
#         dim_data del_noSample del_abond<0.01% del_prev<0.03% del_taxaType dim_dataClean
# OTU         1873            0             261            924            9           933
# Samples      440            0               0              0            0           440
```


```{r}
deleted_taxa_SGB <- as.vector(tax_table(cleanUp2$deleted_taxa)[, "Species"])
table(tax_table(cleanUp2$deleted_taxa)[, "Kingdom"])
 # Archaea Bacteria 
 #       9      931 
```

```{r}
sum(euk %in% deleted_taxa_sp) # 2
sum(euk %in% deleted_taxa_SGB) # 0
length(deleted_taxa_sp) # 679
sum(deleted_taxa_sp %in% deleted_taxa_SGB) # 677 the same species are deleted in sp and SGB
```


```{r message=FALSE}
physeq_rel_SGB_Clean1 = prune_taxa(taxa_sums(cleanUp1$clean_subset) > 0,  cleanUp1$clean_subset)
physeq_rel_SGB_Clean2 = prune_taxa(taxa_sums(cleanUp2$clean_subset) > 0,  cleanUp2$clean_subset)
```


# Normalization 

## Check total number of reads

```{r fig.height=7, fig.width=12, message=FALSE}
phy_obj <- list(physeq_rel_sp=physeq_rel_sp, physeq_rel_sp_Clean1=physeq_rel_sp_Clean1,
                physeq_rel_sp_Clean2=physeq_rel_sp_Clean2, 
                physeq_rel_SGB=physeq_rel_SGB, physeq_rel_SGB_Clean1=physeq_rel_SGB_Clean1,
                physeq_rel_SGB_Clean2=physeq_rel_SGB_Clean2)

pp_list <- list()
for(i in seq(length(phy_obj))){
  phy_obj_i <-phy_obj[[i]]
  sam <- as(sample_data(phy_obj_i), "data.frame")
  sam$sum <- sample_sums(phy_obj_i)
  sam$mean <-colMeans(otu_table(phy_obj_i))
  
  pp_list[[i]]  <- ggplot(sam, aes(x = sum)) + 
    geom_histogram(color = "black", fill = "blue")+  ggtitle("Distribution of sequencing depth") +
    ggtitle(names(phy_obj[i]))
}
wrap_plots(pp_list[1:6], nrow=2) 
```


## Normalization

No need

Not recommended to renormalize after filtering: https://forum.biobakery.org/t/filter-normalize-order-and-comments-on-tutorial/3484

```{r message=FALSE}
# replace physeq_rel_sp and physeq_rel_SBG by the version with only bacterial taxa
physeq_rel_spR = transform_sample_counts(physeq_rel_sp_Clean1, function(x) 100 * x/sum(x))
physeq_rel_SGBR = transform_sample_counts(physeq_rel_SGB_Clean1, function(x) 100 * x/sum(x))

# physeq_rel_sp_Clean2R = transform_sample_counts(physeq_rel_sp_Clean2, function(x) 100 * x/sum(x))
# physeq_rel_SGB_Clean2R = transform_sample_counts(physeq_rel_SGB_Clean2, function(x) 100 * x/sum(x))
```


```{r message=FALSE}
phy_obj <- list(physeq_rel_spR=physeq_rel_spR, physeq_rel_sp_Clean2R=physeq_rel_sp_Clean2R,
                physeq_rel_SGBR=physeq_rel_SGBR, physeq_rel_SGB_Clean2R=physeq_rel_SGB_Clean2R)
for(i in seq(length(phy_obj))){
  phy_obj_i <-phy_obj[[i]]
  print(summary(sample_sums(phy_obj_i)))
}
```

# Tree Summarized Experiment

```{r}
prepareTSE <- function(phy_obj){
  
  tse <- makeTreeSummarizedExperimentFromPhyloseq(phy_obj)
  
  # prepare assays
  mat <- assay(tse, "counts") 
  colsums <- colSums2(mat, na.rm = TRUE)
  if(round(max(colsums)-min(colsums), 3) != 0  ){
    tse <- transformSamples(tse, method = "relabundance") 
    assays(tse)[["relabundance"]] <- assays(tse)[["relabundance"]] *100
    assays(tse)[["inputData"]] <- assays(tse)[["counts"]]  
  }else{
    assays(tse)[["relabundance"]] <- assays(tse)[["counts"]]
  }
  assays(tse)[["counts"]] <- NULL
  
  
  if(is.null(rowTree(tse))){
    cat("Add tree \n")
    tse <- addTaxonomyTree(tse)
  }
  
  ## Add log10 transformation
  tse <- transformFeatures(tse, method = "log10", name = "log10", 
                           pseudocount = 1, assay_name="relabundance")
  
  ## Add CLR transformation
  mat <- assay(tse, "relabundance") 
  pseudo <- min(mat[mat>0])
  tse <- transformSamples(x = tse, assay_name = "relabundance", method = "clr", 
                            pseudocount = pseudo, name = "clr")
    
  
  return(tse)
}
```



```{r message=FALSE}
### Unfiltered data (except archea and eukaryote)
tse <- prepareTSE(physeq_rel_SGBR)
tse_sp <- prepareTSE(physeq_rel_spR)
altExp(tse, "Species") <- tse_sp
```

```{r message=FALSE}
### filtered data 
tse_clean <- prepareTSE(physeq_rel_SGB_Clean2)
tse_sp_clean <- prepareTSE(physeq_rel_sp_Clean2)
altExp(tse_clean, "Species") <- tse_sp_clean
```


# Saving R object

```{r}
save(physeq_rel_spR, physeq_rel_SGBR, tse, 
     file=paste(dir1, "Taxonomy/Data_cleaned_ALL_unfiltered.Rdata", sep=""))

save(physeq_rel_sp_Clean2, physeq_rel_SGB_Clean2, tse_clean, 
     file=paste(dir1, "Taxonomy/Data_cleaned_ALL_filtered.Rdata", sep=""))
```



