---
title: "Analysis of Blood markers"
output: 
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
---



# Loading and formating
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
library(kableExtra)
library(gtsummary)
library(crosstable)
library(data.table)
library(summarytools)
library(skimr)
fix_windows_histograms()
library(lme4)
library(emmeans)
library(gridExtra)
library(tidyr)
library(flextable)
library(circlize)
library(ComplexHeatmap)
library(VennDiagram)
library(RColorBrewer)
library(DHARMa)
library(redres)
library(effects)
library(car)



dir <- "GitHub_repository/IPE_yogurt_rolledOat_microbiome/"
path_analysis <- paste(dir, "DataAnalysis/Results_Main_Analysis/", sep="")


source(file = paste(dir, "R functions/Utils.R", sep=""))
source(file = paste(dir, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir, "R functions/Utils_crossOver.R", sep=""))
source(file = paste(dir, "R functions/code_utils.R", sep=""))

cols <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", 
          Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")
cols_Inter_treat <- cols
```


```{r}
# load(file=paste(dir, "DataPreparation/BloodMarkers/BloodData_Zentrallabor.Rdata", sep=""))
# load(file=paste(dir, "DataPreparation/BloodMarkers/BloodData_Boston.Rdata", sep=""))
load(file=paste(dir, "DataPreparation/BloodMarkers/BloodData_AllMarkers.Rdata", sep="")) # created in Data preparation
load(file=paste(dir, "DataPreparation/Metadata/Info_participant.Rdata", sep=""))
```


```{r paged.print=FALSE}
head(DATA_Boston)
```

```{r paged.print=FALSE}
head(data)
```

```{r paged.print=FALSE}
head(PartiInfo)
```

# Data preparation
```{r}
data_merged <-  merge(DATA_Boston, data[,-c(1,2)], by=c("Participant_ID", "Timepoint"), all=TRUE)
data_merged <-  merge(data_merged, PartiInfo[,c(1:3,6,11,9)], by="Participant_ID", all=TRUE)
```

```{r}
add_meta2 <- function(metadata){
  metadata <- as.data.frame(metadata)
  metadata$Treatment <- case_when(metadata$Timepoint %in% c("T1", "T3") ~ "Before",
                                  metadata$Timepoint %in% c("T2", "T4") ~ "After",
                                  TRUE ~ NA_character_)
  metadata$Intervention <- case_when(metadata$Timepoint %in% c("T1", "T2") & metadata$Gruppe == "A" ~ "Yogurt",
                                     metadata$Timepoint %in% c("T3", "T4") & metadata$Gruppe == "A" ~ "YogurtOatmeal",
                                     metadata$Timepoint %in% c("T1", "T2") & metadata$Gruppe == "B" ~ "YogurtOatmeal",
                                     metadata$Timepoint %in% c("T3", "T4") & metadata$Gruppe == "B" ~ "Yogurt",
                                     TRUE ~ NA_character_)
  
  metadata$Inter_treat <- paste(metadata$Treatment, metadata$Intervention, sep=".")
  metadata$Inter_treat <- factor(metadata$Inter_treat, levels = c("Before.Yogurt", "After.Yogurt", "Before.YogurtOatmeal", "After.YogurtOatmeal"), ordered=T)
  metadata$Treatment <- factor(metadata$Treatment, levels = c("Before", "After"), ordered=T)
  metadata$Timepoint <- factor(metadata$Timepoint, levels = c("T1", "T2", "T3", "T4"), ordered=T)
  metadata$Gruppe <- as.factor(metadata$Gruppe)
  metadata$Participant_ID <- as.factor(metadata$Participant_ID)
  metadata$Geschlecht <- as.factor(metadata$Geschlecht)
  metadata$Intervention <- as.factor(metadata$Intervention)
  
  return(metadata)
}
```



```{r}
data_merged <- add_meta2(data_merged)
```

```{r paged.print=FALSE}
head(data_merged)
```

```{r}
#save(data_merged, file=paste(dir, "DataPreparation/BloodMarkers/BloodData_AllMarkers.Rdata", sep=""))
```


# Data description

```{r}
id_NA <- unique(which(is.na(data_merged[,-c(7, 13:20)]), arr.ind = T)[,"row"])

kbl(data_merged[id_NA,])%>%
  kable_paper("striped", full_width = F, position = "left") 
```
Details of the samples with missing values


## Tables

```{r}
df <- as.data.frame(skim(data_merged[, c(3:6,8,10:12)]))
kbl(df)%>%
  kable_paper("striped", full_width = F, position = "left") 
```


```{r}
var <- colnames(data_merged)[c(3:6,8,10:17)]
tbl_summary(subset(data_merged, Timepoint=="T1")[, var], by=Gruppe, missing ="no") %>% add_p() %>% 
  gtsummary::as_gt()%>% 
  gt::tab_header(title = gt::md("Demographic characteristics and concentrations in the 2 groups at first timepoint"))
```



## Plots

```{r}
measures <- colnames(data_merged)[c(3:6,8,10:12)]
```


### Checking distribution
```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[1]] <- ggplot(data_merged, aes(x=.data[[measures[i]]])) + geom_histogram(color="white") +  theme_classic() + ggtitle("Raw data")
  ll_plot[[2]] <- ll_plot[[1]] + scale_x_continuous(trans = "sqrt")  + ggtitle("square root transformed")
  ll_plot[[3]] <- ll_plot[[1]] + scale_x_continuous(trans = "log2")  + ggtitle("Log2 transformed") 
  ll_plot[[4]] <- ll_plot[[1]] + scale_x_continuous(trans=scales::pseudo_log_trans(base = 2)) + ggtitle("pseudo Log2 transformed") 
  
  print(wrap_plots(ll_plot, nrow=2)  + plot_layout(guides = "collect") + plot_annotation(title=measures[i]))
}
``` 



```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- data_merged[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(x)$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(y))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(x))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
#   measures         shap     shap_log shap_psd_log    shap_sqrt
# 1    TNFR2 4.314743e-15 1.015416e-04 1.015396e-04 8.599940e-10
# 2      IL6 2.610089e-31 9.967513e-10 3.947934e-24 1.891975e-22
# 3     RAGE 4.071492e-03 1.341647e-17 1.345186e-17 9.354760e-05
# 4    8OHdG 5.376825e-07 2.848608e-01 2.782769e-01 1.517566e-03
# 5  zonulin 1.817640e-13 9.465352e-13 2.186900e-12 1.098830e-09
# 6     FRUC 2.909654e-28 1.066586e-22 1.065953e-22 1.390563e-25
# 7     SCRP 1.565336e-39 6.673739e-10 7.788553e-27 1.029653e-30
# 8     SI_H 1.004213e-18 5.766439e-05 3.602187e-04 1.560806e-07
```


```{r paged.print=FALSE}
# transformation for plots
trans <- c("log", "log", "", "", "", "log", "log", "psd_log")
names(trans) <- measures
data.frame(trans)
```

### Blood markers


```{r  fig.height=12, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(data_merged, measures[i], "", x="Treatment", colorInter = T, print_table=F, trans = trans[i]) 
}

wrap_plots(ll_plot, nrow=3)  + plot_layout(guides = "collect") 
```

```{r}
ggsave("ResultsBlood/Blood_byTreat.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect"), width=12, height = 16)
```



```{r fig.height=14, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(data_merged, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```



```{r}
ggsave("ResultsBlood/Blood_byTime.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect"), width=12, height = 14)
```


# Visuals outliers 

## Identification


```{r}
outs <- list()

id_out <- which(data_merged$TNFR2 > 3000 | data_merged$TNFR2 < 1100)
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[1]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[1] <- "TNFR2"

id_out <- which(data_merged$IL6 > 5)
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[2]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[2] <- "IL6"

# id_out <- which(data_merged$RAGE >2100| data_merged$RAGE <200)
# Part_out  <- data_merged[id_out,]$Participant_ID
# outs[[3]] <- subset(data_merged, Participant_ID %in% Part_out)
# names(outs)[3] <- "RAGE"

id_out <- which(data_merged$`8OHdG` >44| data_merged$`8OHdG` <20)
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[3]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[3] <- "8OHdG"


id_out <- which(data_merged$zonulin <10 )
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[4]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[4] <- "zonulin"

id_out <- which(data_merged$FRUC > 320)
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[5]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[5] <- "FRUC"

id_out <- which(data_merged$SCRP > 10 | data_merged$SCRP<0.2)
Part_out  <- data_merged[id_out,]$Participant_ID
outs[[6]] <- subset(data_merged, Participant_ID %in% Part_out)
names(outs)[6] <- "SCRP"
```



```{r fig.height=12, fig.width=15, warning=FALSE}
for(i in seq(length(outs))){
  sub <- outs[[i]]
  title <- names(outs)[i]
  ll_plot <- list()
  for(i in 1:8){
    ll_plot[[i]] <- plot_outliers_path(sub, data_merged, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T)
  }
  
  print(wrap_plots(ll_plot, nrow=3)  + plot_layout(guides = "collect") + 
  plot_annotation(title, theme = theme(plot.title = element_text(size = 22, hjust = 0.5))))
}
```

```{r paged.print=FALSE}
outs2 <- sapply(measures, function(i) as.character(unique(outs[[i]]$Participant_ID)))
repet <- table(unlist(outs2))

df <- data.frame(matrix(ncol = length(measures), nrow = length(names(repet))))
rownames(df) <- names(repet)
colnames(df) <- measures

for(i in seq(length(measures))){
  df[outs2[[i]], measures[i]] <- 1
}
df$total <- rowSums(df, na.rm=T)
df[order(df$total, decreasing=T),]
```

## Filtering

```{r}
data_filt <- data_merged

id <- which(data_filt$Participant_ID == "ID045" & data_filt$Timepoint =="T3" | 
              data_filt$Participant_ID == "ID067" & data_filt$Timepoint =="T2"| 
              data_filt$Participant_ID == "ID014" & data_filt$Timepoint =="T4")
data_filt$TNFR2[id] <- NA

id <- which(data_filt$Participant_ID == "ID067" & data_filt$Timepoint =="T2" | 
              data_filt$Participant_ID == "ID089" & data_filt$Timepoint =="T1"|
              data_filt$Participant_ID == "ID109" & data_filt$Timepoint =="T3")
data_filt$IL6[id] <- NA


id <- which(data_filt$Participant_ID == "ID087" & data_filt$Timepoint =="T2")
data_filt$zonulin[id] <- NA

id <- which(data_filt$Participant_ID == "ID042" & data_filt$Timepoint =="T4" |
              data_filt$Participant_ID == "ID118")
data_filt$FRUC[id] <- NA


id <- which(data_filt$Participant_ID == "ID067" & data_filt$Timepoint =="T2"|
              data_filt$Participant_ID == "ID073" & data_filt$Timepoint =="T1"|
              data_filt$Participant_ID == "ID084" & data_filt$Timepoint =="T4"| 
              data_filt$Participant_ID == "ID089" & data_filt$Timepoint =="T1")
data_filt$SCRP[id] <- NA
```

```{r fig.height=14, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(data_filt, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```


```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[1]] <- ggplot(data_filt, aes(x=.data[[measures[i]]])) + geom_histogram(color="white") +  theme_classic() + ggtitle("Raw data")
  ll_plot[[2]] <- ll_plot[[1]] + scale_x_continuous(trans = "sqrt")  + ggtitle("square root transformed")
  ll_plot[[3]] <- ll_plot[[1]] + scale_x_continuous(trans = "log2")  + ggtitle("Log2 transformed") 
  ll_plot[[4]] <- ll_plot[[1]] + scale_x_continuous(trans=scales::pseudo_log_trans(base = 2)) + ggtitle("pseudo Log2 transformed") 
  
  print(wrap_plots(ll_plot, nrow=2)  + plot_layout(guides = "collect") + plot_annotation(title=measures[i]))
}
``` 

```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- data_filt[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(x)$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(y))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(x))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
#   measures         shap     shap_log shap_psd_log    shap_sqrt
# 1    TNFR2 4.230890e-09 3.786570e-01 3.786548e-01 2.649541e-04
# 2      IL6 9.636421e-30 7.241718e-08 1.154664e-22 1.929333e-20
# 3     RAGE 4.071492e-03 1.341647e-17 1.345186e-17 9.354760e-05
# 4    8OHdG 5.376825e-07 2.848608e-01 2.782769e-01 1.517566e-03
# 5  zonulin 4.103556e-14 9.269206e-04 8.615121e-04 6.162446e-09
# 6     FRUC 3.734990e-06 1.582773e-03 1.582326e-03 9.677682e-05
# 7     SCRP 1.805087e-33 1.613647e-06 1.032367e-23 1.250076e-22
# 8     SI_H 1.004213e-18 5.766439e-05 3.602187e-04 1.560806e-07
```


```{r paged.print=FALSE}
trans_filt <- c("", "log", "", "", "", "", "log", "psd_log")
names(trans_filt) <- measures
data.frame(trans_filt)
```


# Mixed model 


After checking residuals:


* FRUC: 
  * remove outliers points from ID118
  * -> not perfect fit for Alter, negative correlation with BMI: try box-cox transformation
  * -> better general fit
* IL6: 
  * heteroscedasticity, use box-cox transformation
  * -> overfitting for high values, positive correlation with Alter and BMI 
* 8OHdG: 
  * small heteroscedasticity, use box-cox transformation
  * -> not perfect fit for BMI, positive correlation with BMI 
* RAGE: 
  * heavy-tailed, small heteroscedasticity, use box-cox transformation
  * -> overfitting for high values of Alter, negative correlation with Alter and BMI 
* SCRP: 
  * use box-cox transformation
  * -> overfitting for Alter, positive correlation with Alter and BMI  
* SI-H: heteroscedasticity
* TNFR2: 
  * heteroscedasticity, use box-cox transformation
  * -> overfitting for Alter, positive correlation with Alter and BMI 
* zonulin: 
  * remove outlier point ID087_T2
  * -> heteroscedasticity: try box-cox transformation without removing point
  * -> not perfect general fit and deviation uniformity for C1=0, positive correlation with BMI: keep box-cox, remove point again 
  * -> better general fit


3 participants (ID071, ID090, ID113) are removed of the analysis as we do not have their BMI information.

```{r paged.print=FALSE}
DF <- data_merged

# remove influential points
id <- which(DF$Participant_ID == "ID118")
DF$FRUC[id] <- NA

id <- which(DF$Participant_ID == "ID087" & DF$Timepoint =="T2")
DF$zonulin[id] <- NA

id <- which(DF$Participant_ID == "ID067" & DF$Timepoint =="T2")
DF$SCRP[id] <- NA

id <- which(DF$Participant_ID == "ID109" & DF$Timepoint =="T3")
DF$IL6[id] <- NA

# transformation
trans_analysis <- c("bc", "bc", "bc", "bc", "bc", "bc", "bc", "psd_log")
names(trans_analysis) <- measures
data.frame(trans_analysis)
```

```{r}
DF <- dplyr::rename(DF, m_8OHdG=`8OHdG`)
names(trans_analysis)[which(names(trans_analysis) == "8OHdG")] <- "m_8OHdG"
DF_subs <-  prepare_DATA(DF, trans_analysis)
measures_Trans <- sapply(names(trans_analysis), function(x) if(trans_analysis[[x]] !="") paste(x, trans_analysis[[x]], sep="_") else x)
```



```{r}
#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
```

```{r}
sig_threshold <- 0.05
wd0 <- paste(path_analysis, "ResultsBlood/", sep="")
setwd(wd0)

run_uniMarker_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work,
               compar_var=compar_var, var_check=NULL, 
               savingName=paste("lmer", compa, sep="_"), 
               sig_threshold=sig_threshold, check_mod=T, save_mod=F,  
               path=paste("lmer_Trans/", compa, "/", sep=""),
               verbose=F)
```


