---
title: "Analysis of functional KEGG"
output: 
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 4
---


<style>
.main-container { width: 1200px; max-width:2800px;}
</style>





# Loading

```{r message=FALSE}
library(phyloseq)
library(ggplot2)
library(plyr)
library(ape)
library(knitr)
library(vegan)
library(tidyr)
library(gridExtra)
# library(lemon)
library(RColorBrewer)
library(dplyr)
library(cowplot)
library(scater)
library(ggnewscale)
library(mia)
library(miaViz)
library(kableExtra)
library(patchwork)
library(microViz)
library(lme4)
library(emmeans)
library(flextable)
library(ggpubr)
library(reshape)
library(tictoc)
```


```{r message=FALSE}
dir0 <- "Q:/IPE-A-Epidemiologie/Virginie_Stanislas/Projects/"
dir <- "Q:/IPE-P-Joghurtstudie/Joghurt und Haferflocken/Analysis/"
path_analysis <- paste(dir, "DataAnalysis/Results_Complete_Analysis/ResultsFunctional/On_KEGG/", sep="")

source(file = paste(dir, "Utils.R", sep=""))
source(file = paste(dir0, "R functions/Utils_metagenomics.R", sep=""))
source(file = paste(dir0, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir0, "R functions/Utils_crossOver.R", sep=""))
source(file = paste(dir0, "R functions/code_utils.R", sep=""))
```


```{r}
### Load data with already ordination (saved at the end of ordination section)
### contains tse, tse_KEGKEGGnly
load(paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
```


```{r message=FALSE}
load(paste(dir, "DataPreparation/Microbiome_DATA_Kun_Huang/Functional/TSE_Physeq_KEGG.Rdata", sep=""))

tse0 <- tse_KEGG  
physeq0 <- physeq_KEGG 

# Remove 2 samples from ID016 not correctly prepared by the participant
tse <- tse0[, c(!(tse0$Sample_ID  %in% c("JH379", "JH380")))]
physeq <- ps_filter(physeq0, !(Sample_ID %in% c("JH379", "JH380"))) 

tse_KEGGOnly <- tse[!(rowData(tse)$KEGG  %in% c("UNMAPPED", "UNGROUPED")),] #  965 438
physeq_KEGGOnly <- subset_taxa(physeq, !(KEGG  %in% c("UNMAPPED", "UNGROUPED"))) # 965  438 

rm(tse_KEGG, tse0, physeq_KEGG, physeq0)
```


```{r message=FALSE}
# removed ordered factor to allow some plotting functions to work properly
physeq_KEGGOnly_noFact <-  physeq_KEGGOnly
sample_data(physeq_KEGGOnly_noFact)$Inter_treat <- as.character(sample_data(physeq_KEGGOnly_noFact)$Inter_treat )
sample_data(physeq_KEGGOnly_noFact)$Inter_treat <- as.factor(sample_data(physeq_KEGGOnly_noFact)$Inter_treat )
```


```{r}
# proband_colors <- brewer.pal(n = length(unique(sample_data(mydata.psq)$Proband)), name = "Set1")
# names(proband_colors) <- unique(sample_data(mydata.psq)$Proband)

cols <- c("#017b74", "#73ffec", "#ad0053", "#ffcacf", "#ff8d3f", "#ffdf91", "#00c14e","#abffc2", "#6042ed", "#90b9ff", "#700054","#ff8bfc", "#3d8b00", "#bcff7c", "#8b3700", "#ffba91", "#ff4e2a", "#000e26", "#71a800", "#577eff", "#02d0d3", "#d00039", "#dba500", "#ac72ff", "#e0c9ff", "#002e66", "#006dbd", "#02a0c8", "#db37ec",  "#e5ffd5", "#004a20", "#637500", "#ec00bf", "#434d00","#ff77a8", "#002f17", "#f1ff4b", "#533500", "#7ea2ff", "#000d4e", "#ff41a0", "#003636", "#271800", "#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#A65628", "#F781BF", "#FFFF00FF", "#999999")

cols_Inter_treat <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")

cols_Participant_Id <- rep(cols, 3)
names(cols_Participant_Id) <- unique(colData(tse)[, "Participant_ID"])
```


# Exploration

## Prevalence x Abundance

**prevalence**: frequency of samples where the taxa was detected

**abundance**: number of reads assigned to a taxa (on average or in total)




```{r}
rowData(tse)$prevalence <- getPrevalence(tse, detection = 0, sort = FALSE, assay.type = "CPM", as_relative = TRUE)
rowData(tse)$abundance_sum <- rowSums(assays(tse)[["CPM"]])
rowData(tse)$abundance_mean <- rowMeans(assays(tse)[["CPM"]])
rowData(tse)$abundance_max <- apply(assays(tse)[["CPM"]], 1, max)

rowData(tse_KEGGOnly)$prevalence <- getPrevalence(tse_KEGGOnly, detection = 0, sort = FALSE, assay.type = "CPM", as_relative = TRUE)
rowData(tse_KEGGOnly)$abundance_sum <- rowSums(assays(tse_KEGGOnly)[["CPM"]])
rowData(tse_KEGGOnly)$abundance_mean <- rowMeans(assays(tse_KEGGOnly)[["CPM"]])
rowData(tse_KEGGOnly)$abundance_max <- apply(assays(tse_KEGGOnly)[["CPM"]], 1, max)
```


```{r fig.height=7, fig.width=10}
p1 <- ggplot(as.data.frame(rowData(tse)), aes(x=abundance_sum)) + geom_histogram(bins=30, color="white") + scale_x_log10()
p2 <- ggplot(as.data.frame(rowData(tse)), aes(x=abundance_max)) + geom_histogram(bins=30, color="white") + scale_x_log10()
p3 <- ggplot(as.data.frame(rowData(tse)), aes(x=prevalence)) + geom_histogram(bins=30, color="white")
wrap_plots(p1, p2, p3, ncol=2)
```





```{r fig.height=7, fig.width=9}
## Prevalence x abundance on the complete species data
ggplot(as.data.frame(rowData(tse)), aes(abundance_sum, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(labels = scales::label_number(), name = "Total Abundance") +
  scale_y_continuous(
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * ncol(tse), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )) +
  ggrepel::geom_text_repel(
    data = function(df) filter(df, abundance_sum > 300 | prevalence > 0.6),
  mapping = aes(label = KEGG), size = 2.5, min.segment.length = 0)+
   theme_bw()

ggplot(as.data.frame(rowData(tse_KEGGOnly)), aes(abundance_sum, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(labels = scales::label_number(), name = "Total Abundance") +
  scale_y_continuous(
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * ncol(tse), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )) +
  ggrepel::geom_text_repel(
    data = function(df) filter(df, abundance_sum > 300 | prevalence > 0.6),
  mapping = aes(label = KEGG), size = 2.5, min.segment.length = 0)+
   theme_bw()
```



```{r fig.height=10, fig.width=12}
## Prevalence x abundance on the complete species data - axis log transformed
ggplot(as.data.frame(rowData(tse)), aes(abundance_max, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), labels = scales::label_number(), name = "Total Abundance") +
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * ncol(tse), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )) +
  ggrepel::geom_text_repel(
  mapping = aes(label = KEGG), size = 2.5, min.segment.length = 0)+
   theme_bw()

ggplot(as.data.frame(rowData(tse_KEGGOnly)), aes(abundance_max, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), labels = scales::label_number(), name = "Total Abundance") +
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * ncol(tse), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )) +
  ggrepel::geom_text_repel(
  mapping = aes(label = KEGG), size = 2.5, min.segment.length = 0)+
   theme_bw()
```

```{r paged.print=FALSE}
abundance_CPM <- rowSums(assays(tse)[["CPM"]])
abundance_CPM <- abundance_CPM[order(abundance_CPM, decreasing = T)]
head(abundance_CPM, 15)
```

```{r}
plot(abundance_CPM[c(2:500)])
```

```{r paged.print=FALSE}
abundance_CPM2 <- rowSums(assays(tse_KEGGOnly)[["CPM"]])
abundance_CPM2 <- abundance_CPM2[order(abundance_CPM2, decreasing = T)]
head(abundance_CPM2, 15)
```

```{r}
plot(abundance_CPM2[c(2:500)])
```

# Ordination



- **Distance/dissimilarity measures** 
  - **`r colorize("Bray-curtis", "#02a0c8")`**: take into account the abundance $BC_{ij}= 1 - \frac{2C_{ij}}{S_i + S_j}$ with $S_i$ : Total number of species in sample $i$, $S_j$ : Total number of species in sample $j$ and $C_{ij}$ : Sum of lesser value for those species common in both samples
  - **`r colorize("Unifrac", "#02a0c8")`**:  take into account the phylogenetic tree information
      - unweigthed unifrac: Measures the fraction of unique branch length / measures the relative length of those branches that lead exclusively to species present in only one of the two samples with respect to the total length of all branches in the tree.
    - weigthed unifrac: includes information on the relative abundance of each taxa
  - **`r colorize("Jaccard", "#02a0c8")`**: presence / absence, ignores abundance and phylogenetic information
  - **`r colorize("Aitchison", "#02a0c8")`**: take into consideration the compositional structure of the data. Dissimilarity measure calculated as the Euclidean distance between samples after performing a centered log ratio (clr) transformation.
  
- **Ordination/dimention reduction methods** 
  - **`r colorize("PCoA/MDS", "#02a0c8")`**: extension of Principal Components Analysis (PCA),  detect linear trends
  - **`r colorize("NMDS", "#02a0c8")`**: mon-metric multidimensional scaling can detect non-linear patterns
  - **`r colorize("TSNE", "#02a0c8")`**: 
  - **`r colorize("UMAP", "#02a0c8")`**: 
  


```{r}
load(paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
```


```{r}
assay <- "CPM"

### Strange plot when using tse, probably UNMAPPED explain all the variation on first axis??
tse <- runMDS(tse, FUN = vegan::vegdist, method = "bray", name = "PCoA_withUNMAPPED", exprs_values = assay, keep_dist = T)

set.seed(40)
tse <- runNMDS(tse, FUN = vegan::vegdist, method = "bray", name = "NMDS_withUNMAPPED", exprs_values = assay, keep_dist = T)


set.seed(122)
tse <- runTSNE(tse, name = "TSNE_withUNMAPPED", exprs_values=assay)
set.seed(122)
tse <- runUMAP(tse, name = "UMAP_withUNMAPPED", exprs_values=assay)
```


```{r}
tse_KEGGOnly <- runMDS(tse_KEGGOnly, FUN = vegan::vegdist, method = "bray", name = "PCoA_KEGGOnly", exprs_values = assay, keep_dist = T)
set.seed(40)

tse_KEGGOnly <- runNMDS(tse_KEGGOnly, FUN = vegan::vegdist, method = "bray", name = "NMDS_KEGGOnly", exprs_values = assay, keep_dist = T)
set.seed(40)


# Alternative methods to try
# not easy to change the metric for UMAP, no braycurtis dissimilarity in package uwot
set.seed(122)
tse_KEGGOnly <- runTSNE(tse_KEGGOnly, name = "TSNE_KEGGOnly", exprs_values=assay)
set.seed(122)
tse_KEGGOnly <- runUMAP(tse_KEGGOnly, name = "UMAP_KEGGOnly", exprs_values=assay)
```

```{r}
reducedDim(tse, type ="PCoA_KEGGOnly") <- reducedDim(tse_KEGGOnly, "PCoA_KEGGOnly")
reducedDim(tse, type ="NMDS_KEGGOnly") <- reducedDim(tse_KEGGOnly, "NMDS_KEGGOnly")
reducedDim(tse, type ="TSNE_KEGGOnly") <- reducedDim(tse_KEGGOnly, "TSNE_KEGGOnly")
reducedDim(tse, type ="UMAP_KEGGOnly") <- reducedDim(tse_KEGGOnly, "UMAP_KEGGOnly")
```

```{r}
tse_for_ordination <- tse
add_name <- "_withUNMAPPED"

tse_for_ordination <- tse_KEGGOnly
add_name <- "_KEGGOnly"

if(assay == "relabundance") data_assay <- "_rel" 
if(assay == "CPM") data_assay <- "_cpm"
```


## PCoA Bray-Curtis

Almost the same representation when done on filtered data `tse_sub`.
```{r fig.height=12, fig.width=24, message=FALSE, warning=FALSE, paged.print=FALSE}
pp <- plot_ordination(tse_for_ordination, paste("PCoA", add_name, sep=""),
                      sd_threshold=0.01, data_type="Functional", assay=assay)
e <- attr(reducedDim(tse_for_ordination, paste("PCoA", add_name, sep="")), "eig");
rel_eig <- e/sum(e[e>0])          
x_lab = paste("PCoA 1 (", round(100 * rel_eig[[1]],1), "%", ")", sep = "")
y_lab = paste("PCoA 2 (", round(100 * rel_eig[[2]],1), "%", ")", sep = "")

PartHsd_PCoA_BC <- pp$Participants_high_sd
p1 <- pp$p_gruppe + labs(x=x_lab, y=y_lab)
p2 <- pp$p_topdom
p3 <- pp$p_2ndtopdom
p4 <- pp$p_inter_treat + labs(x=x_lab, y=y_lab)
p5 <- pp$p_part + labs(x=x_lab, y=y_lab)
p6 <- pp$p_hv_part + labs(x=x_lab, y=y_lab)
p7 <- pp$p_total_ab + labs(x=x_lab, y=y_lab)
```
Warning message: "'clr' includes negative values. Agglomeration of it might lead to meaningless values." comes from `plot_ordination()` when identifying the most dominant taxa for a specific rank. We can ignore it as we use 'relabudance' assays for identifying dominant taxa.

```{r fig.height=7, fig.width=10, message=FALSE, paged.print=FALSE}
p2
```

```{r}
KEGG_multi <- names(abundance_CPM)[1:13]

p_list <- list()
for(i in seq(length(KEGG_multi))){
KEGG <- KEGG_multi[i]
p_list[[i]] <- plot_ordination(tse, paste("PCoA", add_name, sep=""), functional_id=KEGG, data_type="Functional", assay=assay, top_dom=F)$p_1spec 
}
```


```{r}
ggsave(paste(path_analysis, "Ordination_plots/PCoA", add_name, data_assay, ".png", sep=""), 
       p1 + p5 + p2+ p4 + p6 + p3+ p7 + plot_layout(guides = 'collect'), 
       width=27, height=20)


ggsave(paste(path_analysis, "Ordination_plots/PCoA", add_name, data_assay, "_withAbundance.png", sep=""), 
       wrap_plots(p_list, ncol=4) , 
       width=30, height=20)
```



## NMDS Bray-Curtis

Almost the same representation when done on filtered data `tse_sub`.
```{r fig.height=12, fig.width=24, message=FALSE, warning=FALSE, paged.print=FALSE}
pp <- plot_ordination(tse_for_ordination, paste("NMDS", add_name, sep=""), 
                      sd_threshold=0.05, data_type="Functional", assay=assay)
PartHsd_NMDS_BC <- pp$Participants_high_sd
p1 <- pp$p_gruppe 
p2 <- pp$p_topdom
p3 <- pp$p_2ndtopdom
p4 <- pp$p_inter_treat 
p5 <- pp$p_part 
p6 <- pp$p_hv_part
p7 <- pp$p_total_ab 
```


```{r fig.height=7, fig.width=10, message=FALSE, paged.print=FALSE}
p2
```

```{r}
KEGG_multi <- names(abundance_CPM)[1:13]

p_list <- list()
for(i in seq(length(KEGG_multi))){
KEGG <- KEGG_multi[i]
p_list[[i]] <- plot_ordination(tse, paste("NMDS", add_name, sep=""), functional_id=KEGG, data_type="Functional", assay=assay, top_dom=F)$p_1spec 
}
```


```{r}
ggsave(paste(path_analysis, "Ordination_plots/NMDS", add_name, data_assay,".png", sep=""), 
       p1 + p5 + p2+ p4 + p6 + p3+ p7 + plot_layout(guides = 'collect'), 
       width=27, height=20)

ggsave(paste(path_analysis, "Ordination_plots/NMDS", add_name, data_assay, "_withAbundance.png", sep=""), 
       wrap_plots(p_list, ncol=4) , 
       width=30, height=20)
```



## TSNE
```{r fig.height=12, fig.width=24, message=FALSE, warning=FALSE, paged.print=FALSE}
pp <- plot_ordination(tse_for_ordination, paste("TSNE", add_name, sep=""),
                      sd_threshold=10, data_type="Functional", assay=assay)
PartHsd_TSNE <- pp$Participants_high_sd
p1 <- pp$p_gruppe 
p2 <- pp$p_topdom
p3 <- pp$p_2ndtopdom
p4 <- pp$p_inter_treat 
p5 <- pp$p_part 
p6 <- pp$p_hv_part
p7 <- pp$p_total_ab 
```


```{r fig.height=7, fig.width=10, message=FALSE, paged.print=FALSE}
p2
```

```{r}
KEGG_multi <- names(abundance_CPM)[1:13]

p_list <- list()
for(i in seq(length(KEGG_multi))){
KEGG <- KEGG_multi[i]
p_list[[i]] <- plot_ordination(tse, paste("TSNE", add_name, sep=""), functional_id=KEGG, data_type="Functional", assay=assay, top_dom=F)$p_1spec 
}
```


```{r}
ggsave(paste(path_analysis, "Ordination_plots/TSNE", add_name, data_assay,".png", sep=""),
       p1 + p5 + p2+ p4 + p6 + p3+ p7 + plot_layout(guides = 'collect'), 
       width=27, height=20)

ggsave(paste(path_analysis, "Ordination_plots/TSNE", add_name, data_assay, "_withAbundance.png", sep=""), 
       wrap_plots(p_list, ncol=4) , 
       width=30, height=20)
```


## UMAP



```{r fig.height=12, fig.width=24, message=FALSE, warning=FALSE}
pp <- plot_ordination(tse_for_ordination, paste("UMAP", add_name, sep=""), sd_threshold=2, data_type="Functional", assay=assay)
PartHsd_UMAP <- pp$Participants_high_sd
p1 <- pp$p_gruppe
p2 <- pp$p_topdom
p3 <- pp$p_2ndtopdom
p4 <- pp$p_inter_treat 
p5 <- pp$p_part 
p6 <- pp$p_hv_part
p7 <- pp$p_total_ab
```


```{r fig.height=7, fig.width=10, message=FALSE, paged.print=FALSE}
p2
```

```{r}
KEGG_multi <- names(abundance_CPM)[1:13]

p_list <- list()
for(i in seq(length(KEGG_multi))){
KEGG <- KEGG_multi[i]
p_list[[i]] <- plot_ordination(tse, paste("UMAP", add_name, sep=""), functional_id=KEGG, data_type="Functional", assay=assay, top_dom=F)$p_1spec 
}
```


```{r}
ggsave(paste(path_analysis, "Ordination_plots/UMAP", add_name, data_assay, ".png", sep=""),
       p1 + p5 + p2+ p4 + p6 + p3+ p7 + plot_layout(guides = 'collect'), 
       width=27, height=20)

ggsave(paste(path_analysis, "Ordination_plots/UMAP", add_name, data_assay, "_withAbundance.png", sep=""), 
       wrap_plots(p_list, ncol=4) , 
       width=30, height=20)
```

  

# Saving tse
```{r}
save(tse, tse_KEGGOnly, file=paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
# data with ordination made on CPM
# To keep same ordination (ran on Computer office)
```


# Bar plots

```{r fig.height=12, fig.width=12, message=FALSE, warning=FALSE}
n_taxa <- 50
cols_mod <- cols
cols_mod[n_taxa+1] <- "lightgrey"

physeq_KEGGOnly_noFact  %>%
  ps_select(Participant_ID, Inter_treat) %>%
  phyloseq::merge_samples(group = "Inter_treat") %>%
  comp_barplot(tax_level = "KEGG", n_taxa = n_taxa, bar_width = 0.9, palette=cols_mod,
               tax_transform_for_plot="identity", bar_outline_colour = NA,
               sample_order=c("Before.Yogurt", "After.Yogurt", "Before.YogurtOatmeal", "After.YogurtOatmeal")) +
  labs(x = NULL, y = NULL) + guides(fill = guide_legend(ncol = 1, reverse=TRUE)) +theme(legend.key.size = unit(0.5, 'cm'))
```





# Beta diversity


```{r message=FALSE}
metadata <- as.data.frame(colData(tse))
assay_t2 <- assay_t <- t(assays(tse)[["CPM"]]) %>% as.data.frame()
KEGG <- colnames(assay_t2)
assay_t2$Sample_ID <- rownames(assay_t2)
assay_t2 <- merge(metadata, assay_t2, by="Sample_ID")

assay_t_Y <- subset(assay_t2,  Intervention== "Yogurt")
assay_t_YO <- subset(assay_t2,  Intervention== "YogurtOatmeal")
assay_t_Y <- assay_t_Y[, KEGG]
assay_t_YO <- assay_t_YO[, KEGG]
```

```{r}
Perct0 <- apply(assay_t, 2, function(x) sum(x==0)/nrow(assay_t)*100)
Perct0_Y <- apply(assay_t_Y, 2, function(x) sum(x==0)/nrow(assay_t_Y)*100)
Perct0_YO <- apply(assay_t_YO, 2, function(x) sum(x==0)/nrow(assay_t_YO)*100)

hist(Perct0)
```


```{r}
KEGG_infos <- data.frame(Perct0, Perct0_Y, Perct0_YO)
KEGG_infos$KEGG <- rownames(KEGG_infos)
KEGG_infos$prev <- 100-KEGG_infos$Perct0
KEGG_infos$prev_Y <- 100-KEGG_infos$Perct0_Y
KEGG_infos$prev_YO <- 100-KEGG_infos$Perct0_YO
KEGG_infos <- KEGG_infos[, c(4,1:3,5:7)]
```

```{r}
# subset(KEGG_infos, Perct0_YO==0 & Perct0>0) # 7   7
# subset(KEGG_infos, Perct0_Y==0 & Perct0>0) # 93  7

KEGG_perct0_0 <- subset(KEGG_infos, Perct0 <= 0)$KEGG
KEGG_perct0_sup0 <- subset(KEGG_infos, Perct0 > 0)$KEGG
KEGG_no_fitted <- KEGG_infos[which(!(KEGG_infos$KEGG %in% KEGG_perct0_0) & !(KEGG_infos$KEGG %in% KEGG_perct0_sup0)), "KEGG"]

KEGG_perct0_0_Y <- subset(KEGG_infos, Perct0_Y <= 0)$KEGG
KEGG_perct0_sup0_Y <- subset(KEGG_infos, Perct0_Y > 0 )$KEGG
KEGG_no_fitted_Y <- KEGG_infos[which(!(KEGG_infos$KEGG %in% KEGG_perct0_0_Y) & !(KEGG_infos$KEGG %in% KEGG_perct0_sup0_Y)), "KEGG"]

KEGG_perct0_0_YO <- subset(KEGG_infos, Perct0_YO <= 0)$KEGG
KEGG_perct0_sup0_YO <- subset(KEGG_infos, Perct0_YO > 0)$KEGG
KEGG_no_fitted_YO <- KEGG_infos[which(!(KEGG_infos$KEGG %in% KEGG_perct0_0_YO) & !(KEGG_infos$KEGG %in% KEGG_perct0_sup0_YO)), "KEGG"]
```


```{r}
length(KEGG_perct0_0) # 62 KEGG have no zero values in both intervention
length(KEGG_perct0_sup0) # 905 KEGG have zero values in at least one of the intervention
length(KEGG_no_fitted)
# [1] 62
# [1] 905
# [1] 0

length(KEGG_perct0_0_Y) # 155 (62 + 93) KEGG have no zero values in Yogurt
length(KEGG_perct0_sup0_Y) # 812 KEGG have zero values in Yogurt
length(KEGG_no_fitted_Y)
# [1] 155
# [1] 812
# [1] 0

length(KEGG_perct0_0_YO)  # 69 (62 + 7) KEGG have no zero values in YogurtOatmeal
length(KEGG_perct0_sup0_YO)  # 898 KEGG have zero values in YogurtOatmeal
length(KEGG_no_fitted_YO)
# [1] 69
# [1] 898
# [1] 0
```

## transformation no covariates


```{r message=FALSE}
assay_t <- t(assays(tse)[["CPM"]]) %>% as.data.frame()
KEGG <- colnames(assay_t)
assay_t$Sample_ID <- rownames(assay_t)
metadata <- as.data.frame(colData(tse))
DF <- merge(metadata, assay_t, by="Sample_ID")
DF_subs <- prepare_DATA(DF, NULL)
```

```{r}
#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"
KEGG_to_use <- KEGG_perct0_0
KEGG_to_use_w0 <- KEGG_perct0_sup0

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"
KEGG_to_use <- KEGG_perct0_0_Y
KEGG_to_use_w0 <- KEGG_perct0_sup0_Y

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
KEGG_to_use <- KEGG_perct0_0_YO
KEGG_to_use_w0 <- KEGG_perct0_sup0_YO
```


### LMER on KEGG without zeroes
```{r}
model <- "lmer_bc"
trans <- ""
sig_threshold <- 0.03
```

```{r}
trans_var <- rep("bc", length(KEGG_to_use))
names(trans_var) <- KEGG_to_use
DF_work_mod <- transform_data(DF_work, trans_var)
measures_Trans <- sapply(names(trans_var), function(x) if(trans_var[[x]] !="") paste(x, trans_var[[x]], sep="_") else x)
```


```{r}
tic()
# lmer
wd0 <- paste(path_analysis, "ResultsBetadiv/", sep="")
setwd(wd0)
run_taxa_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("lmer_bc_", compa, sep=""), sig_threshold=sig_threshold, check_mod=F, save_mod=F,  path=paste("lmer_bc/", compa, "/", sep=""))
toc()
# Model full: 44.03 sec elapsed
# Model Y:  76.57 sec elapsed
# Model YO: 39.26 sec elapsed
```



### ZIG on KEGG with zeroes
```{r}
library(NBZIMM)
```


```{r}
model <- "NBZIMM_asinsqrt"
trans <- ""
sig_threshold <- 0.03
```

```{r}
DF_work_mod=DF_work
DF_work_mod[,KEGG] <- asin(sqrt(DF_work_mod[, KEGG]/1000000)) 
```

```{r}
tic()
# ZIG
wd0 <- paste(path_analysis, "ResultsBetadiv/", sep="")
setwd(wd0)
run_taxa_model(KEGG_to_use_w0, package="NBZIMM", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("NBZIMM_asinsqrt_", compa, sep=""), sig_threshold=sig_threshold, check_mod=F, save_mod=F,  path=paste("NBZIMM_asinsqrt/", compa, "/", sep=""))
toc()
# Model full: 418.33 sec elapsed
# Model Y: 353.39 sec elapsed
# Model YO: 399.87 sec elapsed
```

### ZIG on all KEGG run by Kun (Same results than previous section)


```{r message=FALSE}
assay_t <- t(assays(tse)[["CPM"]]) %>% as.data.frame()
KEGG <- colnames(assay_t)
assay_t$Sample_ID <- rownames(assay_t)
metadata <- as.data.frame(colData(tse))

DF_full <- merge(metadata, assay_t, by="Sample_ID")
DF_Y <- subset(DF_full,  Intervention== "Yogurt")
DF_YO <- subset(DF_full,  Intervention== "YogurtOatmeal")

DF_Y <- droplevels(DF_Y)  
DF_YO <- droplevels(DF_YO)  
```

#### Model full
```{r}
#### Model full
compar_var = "T"
DF_work=DF_full
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"
```

```{r}
DF_work_mod=DF_work
DF_work_mod[, KEGG] <- asin(sqrt(DF_work_mod[, KEGG]/1000000)) 
```

```{r}
### RAN BY KUN
# # ZIG
# wd0 <- paste(path_analysis, "ResultsBetadiv/", sep="")
# setwd(wd0)
# run_taxa_model(KEGG, package="NBZIMM", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("NBZIMM_asinsqrt_", compa, sep=""), check_mod=F, save_mod=F,  path=paste("NBZIMM_asinsqrt_Kun/", compa, "/", sep=""), save_final_files=F)
```

```{r}
model <- "NBZIMM_asinsqrt"

path <- paste(path_analysis, "ResultsBetadiv/NBZIMM_asinsqrt_Kun/", compa, "/",  sep="")

res_all_full <- read.table(file=paste(path, model, "_", compa,"_res_all_results.tsv", sep=""))
res_anov_all_full <- read.table(file=paste(path, model, "_", compa,"_res_anov_results.tsv", sep=""))
PRS_all_full <- read.table(file=paste(path, model, "_", compa,"_PRS_all_results.tsv", sep=""))
```



#### Model Y
```{r}
#### Model Y
compar_var = "B"
DF_work=DF_Y
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"
```


```{r}
DF_work_mod=DF_work
DF_work_mod[,KEGG] <- asin(sqrt(DF_work_mod[, KEGG]/1000000)) 
```


```{r}
### RAN BY KUN
# # ZIG
# wd0 <- paste(path_analysis, "ResultsBetadiv/", sep="")
# setwd(wd0)
# run_taxa_model(KEGG, package="NBZIMM", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("NBZIMM_asinsqrt_", compa, sep=""), check_mod=F, save_mod=F,  path=paste("NBZIMM_asinsqrt_Kun/", compa, "/", sep=""), save_final_files=F)
```

```{r}
model <- "NBZIMM_asinsqrt"

path <- paste(path_analysis, "ResultsBetadiv/NBZIMM_asinsqrt_Kun/", compa, "/", sep="")

res_all_Y <- read.table(file=paste(path, model, "_", compa,"_res_all_results.tsv", sep=""))
res_anov_all_Y <- read.table(file=paste(path, model, "_", compa,"_res_anov_results.tsv", sep=""))
PRS_all_Y <- read.table(file=paste(path, model, "_", compa,"_PRS_all_results.tsv", sep=""))
```





#### Model YO
```{r}
#### Model YO
compar_var = "B"
DF_work=DF_YO
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
```


```{r}
DF_work_mod=DF_work
DF_work_mod[,KEGG] <- asin(sqrt(DF_work_mod[, KEGG]/1000000)) 
```

```{r}
### RAN BY KUN
# # ZIG
# wd0 <- paste(path_analysis, "ResultsBetadiv/", sep="")
# setwd(wd0)
# run_taxa_model(KEGG, package="NBZIMM", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("NBZIMM_asinsqrt_", compa, sep=""), check_mod=F, save_mod=F,  path=paste("NBZIMM_asinsqrt_Kun/", compa, "/", sep=""), save_final_files=F)
```

```{r}
model <- "NBZIMM_asinsqrt"

path <- paste(path_analysis, "ResultsBetadiv/NBZIMM_asinsqrt_Kun/", compa, "/", sep="")

res_all_YO <- read.table(file=paste(path, model, "_", compa,"_res_all_results.tsv", sep=""))
res_anov_all_YO <- read.table(file=paste(path, model, "_", compa,"_res_anov_results.tsv", sep=""))
PRS_all_YO <- read.table(file=paste(path, model, "_", compa,"_PRS_all_results.tsv", sep=""))
```




