---
title: "SCFA analysis"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
---



```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
library(kableExtra)
library(gtsummary)
library(crosstable)
library(data.table)
library(summarytools)
library(skimr)
fix_windows_histograms()
library(lme4)
library(emmeans)
library(gridExtra)
library(tidyr)
library(flextable)
library(circlize)
library(ComplexHeatmap)
library(VennDiagram)
library(RColorBrewer)
library(DHARMa)
library(redres)
library(effects)
library(car)


dir <- "GitHub_repository/IPE_yogurt_rolledOat_microbiome/"
path_analysis <- paste(dir, "DataAnalysis/Results_Main_Analysis/", sep="")

source(file = paste(dir, "R functions/Utils.R", sep=""))
source(file = paste(dir, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir, "R functions/Utils_crossOver.R", sep=""))
source(file = paste(dir, "R functions/code_utils.R", sep=""))

cols <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", 
          Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")
cols_Inter_treat <- cols
```



```{r}
load(file=paste(dir, "DataPreparation/Metabolomic/MetabolomeData.Rdata", sep=""))
```



```{r paged.print=FALSE}
head(DATA)
```


# Data description

```{r}
id_NA <- unique(DATA[is.na(DATA$Results), "Sample_ID"])

kbl(subset(DATA_wide, Sample_ID %in% id_NA))%>%
  kable_paper("striped", full_width = F, position = "left") 
```
Details of the samples with missing values


## Tables

```{r}
df <- as.data.frame(skim(DATA_wide[, c(12:19)]))[, c(2:7, 9, 11, 12)]
kbl(df)%>%
  kable_paper("striped", full_width = F, position = "left") 
```



```{r}
var <- colnames(DATA_wide)[c(6:4,12:19)]
tbl_summary(subset(DATA_wide, Timepoint=="T1")[, var], by=Gruppe, missing ="no") %>% add_p() %>% 
  gtsummary::as_gt()%>% 
  gt::tab_header(title = gt::md("Demographic characteristics and concentrations in the 2 groups at first timepoint"))
```



## Plots
```{r}
measures <- colnames(DATA_wide)[12:19]
```


### Checking distribution
```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[1]] <- ggplot(DATA_wide, aes(x=.data[[measures[i]]])) + geom_histogram(color="white") +  theme_classic() + ggtitle("Raw data")
  ll_plot[[2]] <- ll_plot[[1]] + scale_x_continuous(trans = "sqrt")  + ggtitle("square root transformed")
  ll_plot[[3]] <- ll_plot[[1]] + scale_x_continuous(trans = "log2")  + ggtitle("Log2 transformed") 
  ll_plot[[4]] <- ll_plot[[1]] + scale_x_continuous(trans=scales::pseudo_log_trans(base = 2)) + ggtitle("pseudo Log2 transformed") 
  
  print(wrap_plots(ll_plot, nrow=2)  + plot_layout(guides = "collect") + plot_annotation(title=measures[i]))
}
``` 



```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- DATA_wide[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(DATA_wide[,measures[i]])$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(DATA_wide[,measures[i]]))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(DATA_wide[,measures[i]]))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
#              measures         shap     shap_log shap_psd_log    shap_sqrt
# 1 Methylbutyric_acid2 4.711859e-18 7.712766e-22 8.255604e-13 4.200205e-04
# 2         Acetic_acid 5.950772e-08 3.291859e-06 3.291978e-06 3.044447e-02
# 3        Butyric_acid 3.541571e-14 8.438155e-17 8.834752e-17 2.226244e-01
# 4       Hexanoic_acid 2.113432e-22 1.888815e-14 1.083219e-15 9.264651e-15
# 5     Isobutyric_acid 6.511473e-18 1.044187e-10 1.890671e-10 2.046469e-06
# 6     Isovaleric_acid 2.575904e-18 1.724201e-07 2.255240e-07 4.128361e-06
# 7      Propionic_acid 6.617865e-14 3.811778e-05 3.813919e-05 2.094209e-03
# 8        Valeric_acid 1.109376e-12 4.862326e-16 5.386770e-16 4.643904e-03
```


```{r paged.print=FALSE}
trans <- c("sqrt", "sqrt", "sqrt", "psd_log", "sqrt", "sqrt", "sqrt", "sqrt")
names(trans) <- measures
data.frame(trans)
```

### SCFAs



```{r fig.height=16, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Treatment", colorInter = T, print_table=F, trans = trans[i]) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```

```{r}
ggsave("ResultsMetabolomic/SCFA_byTreat.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") , width=15, height = 14)
```



```{r fig.height=32, fig.width=9, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=8)  + plot_layout(guides = "collect") 
```



```{r}
ggsave("ResultsMetabolomic/SCFA_byTime.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") , width=18, height = 16)
```





# Visuals outliers 

## Identification


```{r}
outs <- list()

id_out <- which(DATA_wide$Methylbutyric_acid2 > 320 | DATA_wide$Methylbutyric_acid2 < 7)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[1]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[1] <- "Methylbutyric_acid2"


id_out <- which(DATA_wide$Acetic_acid > 7100 | DATA_wide$Acetic_acid < 200)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[2]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[2] <- "Acetic_acid"


id_out <- which(DATA_wide$Butyric_acid > 4500 | DATA_wide$Butyric_acid < 70)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[3]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[3] <- "Butyric_acid"


id_out <- which(DATA_wide$Hexanoic_acid > 900 | DATA_wide$Hexanoic_acid < 0.5)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[4]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[4] <- "Hexanoic_acid"


id_out <- which(DATA_wide$Isobutyric_acid > 450 | DATA_wide$Isobutyric_acid < 15)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[5]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[5] <- "Isobutyric_acid"


id_out <- which(DATA_wide$Isovaleric_acid > 450 | DATA_wide$Isovaleric_acid < 10)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[6]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[6] <- "Isovaleric_acid"


id_out <- which(DATA_wide$Propionic_acid > 3000 | DATA_wide$Propionic_acid < 50)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[7]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[7] <- "Propionic_acid"


id_out <- which(DATA_wide$Valeric_acid > 620 | DATA_wide$Valeric_acid < 20)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[8]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[8] <- "Valeric_acid"
```



```{r fig.height=15, fig.width=15, warning=FALSE}
for(i in seq(length(outs))){
  sub <- outs[[i]]
  title <- names(outs)[i]
  ll_plot <- list()
  for(i in 1:8){
    ll_plot[[i]] <- plot_outliers_path(sub, DATA_wide, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T)
  }
  
  print(wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") + 
  plot_annotation(title, theme = theme(plot.title = element_text(size = 22, hjust = 0.5))))
}
```


```{r paged.print=FALSE}
outs2 <- sapply(measures, function(i) as.character(unique(outs[[i]]$Participant_ID)))
repet <- table(unlist(outs2))

df <- data.frame(matrix(ncol = length(measures), nrow = length(names(repet))))
rownames(df) <- names(repet)
colnames(df) <- measures

for(i in seq(length(measures))){
  df[outs2[[i]], measures[i]] <- 1
}
df$total <- rowSums(df, na.rm=T)
df[order(df$total, decreasing=T),]
```


## Filtering

```{r}
data_filt <- DATA_wide

id <- which(data_filt$Participant_ID == "ID099")
data_filt$Methylbutyric_acid2[id] <- NA
data_filt$Isobutyric_acid[id] <- NA
data_filt$Isovaleric_acid[id] <- NA

id <- which(data_filt$Participant_ID == "ID077" & data_filt$Timepoint =="T1"|
              data_filt$Participant_ID == "ID119" & data_filt$Timepoint =="T4"|
              data_filt$Participant_ID == "ID017" & data_filt$Timepoint =="T4")
data_filt$Acetic_acid[id] <- NA


id <- which(data_filt$Participant_ID == "ID077" & data_filt$Timepoint =="T1")
data_filt$Butyric_acid[id] <- NA
```

```{r fig.height=14, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(data_filt, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```



```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- data_filt[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(x)$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(y))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(x))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
#              measures         shap     shap_log shap_psd_log    shap_sqrt
# 1 Methylbutyric_acid2 1.209058e-14 2.778551e-22 1.538899e-13 1.006943e-01
# 2         Acetic_acid 7.174755e-08 6.022639e-06 6.022801e-06 1.274926e-02
# 3        Butyric_acid 3.069579e-14 7.716522e-16 8.093656e-16 1.998496e-01
# 4       Hexanoic_acid 2.113432e-22 1.888815e-14 1.083219e-15 9.264651e-15
# 5     Isobutyric_acid 1.667079e-15 2.925885e-11 5.253142e-11 2.244956e-04
# 6     Isovaleric_acid 9.225688e-17 4.607103e-08 5.998526e-08 9.426451e-05
# 7      Propionic_acid 6.617865e-14 3.811778e-05 3.813919e-05 2.094209e-03
# 8        Valeric_acid 1.109376e-12 4.862326e-16 5.386770e-16 4.643904e-03
```

```{r paged.print=FALSE}
trans_filt <- c("sqrt", "sqrt", "sqrt", "psd_log", "sqrt", "sqrt", "sqrt", "sqrt")
names(trans_filt) <- measures
data.frame(trans_filt)
```





# Mixed model


After checking residuals:

* Acetic_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> overfitting for Alter, negative correlation with Alter and positive correlation with BMI  
* Butyric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter
* Hexanoic_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> bimodality in the data
* Isobutyric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Isovaleric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Methylbutyric_acid2: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Propionic_acid: 
  * heteroscedasticity, use box-cox transformation
  * no general good fit for high value, non homogeneity of variance for Geschlecht, cubic relation with Alter and BMI, negative correlation with Alter and positive correlation with BMI 
* Valeric_acid:
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, overfitting ALter, BMI, negative correlation with Alter and positive correlation with BMI 
  
  
  

  
```{r paged.print=FALSE}
# transformation
trans_analysis <- c("bc", "bc", "bc", "bc", "bc", "bc", "bc", "bc")
names(trans_analysis) <- measures
data.frame(trans_analysis)
```

```{r}
DF_subs <-  prepare_DATA(DATA_wide, trans_analysis)
measures_Trans <- sapply(names(trans_analysis), function(x) if(trans_analysis[[x]] !="") paste(x, trans_analysis[[x]], sep="_") else x)
```



```{r}
#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
```

```{r}
sig_threshold <- 0.05
wd0 <- paste(path_analysis, "ResultsMetabolomic/", sep="")
setwd(wd0)

run_uniMarker_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work,
               compar_var=compar_var, var_check=NULL, 
               savingName=paste("lmer", compa, sep="_"), 
               sig_threshold=sig_threshold, check_mod=T, save_mod=F,  
               path=paste("lmer_Trans/", compa, "/", sep=""),
               verbose=F)
```
