---
title: "SCFA analysis"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
---



```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
library(kableExtra)
library(gtsummary)
library(crosstable)
library(data.table)
library(summarytools)
library(skimr)
fix_windows_histograms()
library(lme4)
library(emmeans)
library(gridExtra)
library(tidyr)
library(flextable)
library(circlize)
library(ComplexHeatmap)
library(VennDiagram)
library(RColorBrewer)
library(DHARMa)
library(redres)
library(effects)
library(car)

dir0 <- "Q:/IPE-A-Epidemiologie/Virginie_Stanislas/Projects/"
dir <- "Q:/IPE-P-Joghurtstudie/Joghurt und Haferflocken/Analysis/"
path_analysis <- paste(dir, "DataAnalysis/Results_Complete_Analysis/", sep="")


source(file = paste(dir, "Utils.R", sep=""))
source(file = paste(dir0, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir0, "R functions/Utils_crossOver.R", sep=""))
source(file = paste(dir0, "R functions/code_utils.R", sep=""))

cols <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", 
          Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")
cols_Inter_treat <- cols
```



```{r}
load(file=paste(dir, "DataPreparation/Metabolomic/MetabolomeData.Rdata", sep=""))
```



```{r paged.print=FALSE}
head(DATA)
```


# Data description

```{r}
id_NA <- unique(DATA[is.na(DATA$Results), "Sample_ID"])

kbl(subset(DATA_wide, Sample_ID %in% id_NA))%>%
  kable_paper("striped", full_width = F, position = "left") 
```
Details of the samples with missing values


## Tables

```{r}
df <- as.data.frame(skim(DATA_wide[, c(12:19)]))[, c(2:7, 9, 11, 12)]
kbl(df)%>%
  kable_paper("striped", full_width = F, position = "left") 
```

```{r}
var <- colnames(DATA_wide)[c(3:4,12:19)]
tbl_summary(DATA_wide[, var], by=Gruppe, missing ="no") %>% #add_p() %>% 
  gtsummary::as_gt()%>% 
  gt::tab_header(title = gt::md("Comparisons between the two groups"))
```



```{r}
var <- colnames(DATA_wide)[c(6:4,12:19)]
tbl_summary(subset(DATA_wide, Timepoint=="T1")[, var], by=Gruppe, missing ="no") %>% add_p() %>% 
  gtsummary::as_gt()%>% 
  gt::tab_header(title = gt::md("Demographic characteristics and concentrations in the 2 groups at first timepoint"))
```



## Plots
```{r}
measures <- colnames(DATA_wide)[12:19]
```


### Checking distribution
```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[1]] <- ggplot(DATA_wide, aes(x=.data[[measures[i]]])) + geom_histogram(color="white") +  theme_classic() + ggtitle("Raw data")
  ll_plot[[2]] <- ll_plot[[1]] + scale_x_continuous(trans = "sqrt")  + ggtitle("square root transformed")
  ll_plot[[3]] <- ll_plot[[1]] + scale_x_continuous(trans = "log2")  + ggtitle("Log2 transformed") 
  ll_plot[[4]] <- ll_plot[[1]] + scale_x_continuous(trans=scales::pseudo_log_trans(base = 2)) + ggtitle("pseudo Log2 transformed") 
  
  print(wrap_plots(ll_plot, nrow=2)  + plot_layout(guides = "collect") + plot_annotation(title=measures[i]))
}
``` 

```{r eval=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
#qqplot to check normal distribution
ll_plot <- list()
ll_plot2 <- list()
ll_plot3 <- list()
for(i in 1:8){
  ll_plot[[i]] <- ggqqplot(DATA_wide[,measures[i]])
  ll_plot2[[i]] <- ggqqplot(log2(DATA_wide[,measures[i]]))
  ll_plot3[[i]] <- ggqqplot(sqrt(DATA_wide[,measures[i]]))
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") + plot_annotation(title="Raw data")
wrap_plots(ll_plot2, nrow=4)  + plot_layout(guides = "collect") + plot_annotation(title="Log transformed")
wrap_plots(ll_plot3, nrow=4)  + plot_layout(guides = "collect") + plot_annotation(title="sqrt transformed")
``` 


```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- DATA_wide[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(DATA_wide[,measures[i]])$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(DATA_wide[,measures[i]]))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(DATA_wide[,measures[i]]))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
```


```{r paged.print=FALSE}
trans <- c("sqrt", "sqrt", "sqrt", "psd_log", "sqrt", "sqrt", "sqrt", "sqrt")
names(trans) <- measures
data.frame(trans)
```

### SCFAs
```{r fig.height=16, fig.width=12, warning=FALSE}
ll_plot <- list()
printMean <- c(rep(T, 3), F, rep(T, 4))
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Treatment", colorInter = T, print_table=F, trans = trans[i], printMean=printMean[i], line_diff = F) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```



```{r fig.height=16, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Treatment", colorInter = T, print_table=F, trans = trans[i]) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```

```{r}
ggsave("ResultsMetabolomic/SCFA_byTreat.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") , width=15, height = 14)
```



```{r fig.height=32, fig.width=9, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=8)  + plot_layout(guides = "collect") 
```



```{r}
ggsave("ResultsMetabolomic/SCFA_byTime.png", wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") , width=18, height = 16)
```





# Visuals outliers 

## Identification


```{r}
outs <- list()

id_out <- which(DATA_wide$Methylbutyric_acid2 > 320 | DATA_wide$Methylbutyric_acid2 < 7)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[1]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[1] <- "Methylbutyric_acid2"


id_out <- which(DATA_wide$Acetic_acid > 7100 | DATA_wide$Acetic_acid < 200)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[2]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[2] <- "Acetic_acid"


id_out <- which(DATA_wide$Butyric_acid > 4500 | DATA_wide$Butyric_acid < 70)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[3]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[3] <- "Butyric_acid"


id_out <- which(DATA_wide$Hexanoic_acid > 900 | DATA_wide$Hexanoic_acid < 0.5)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[4]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[4] <- "Hexanoic_acid"


id_out <- which(DATA_wide$Isobutyric_acid > 450 | DATA_wide$Isobutyric_acid < 15)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[5]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[5] <- "Isobutyric_acid"


id_out <- which(DATA_wide$Isovaleric_acid > 450 | DATA_wide$Isovaleric_acid < 10)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[6]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[6] <- "Isovaleric_acid"


id_out <- which(DATA_wide$Propionic_acid > 3000 | DATA_wide$Propionic_acid < 50)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[7]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[7] <- "Propionic_acid"


id_out <- which(DATA_wide$Valeric_acid > 620 | DATA_wide$Valeric_acid < 20)
Part_out  <- DATA_wide[id_out,]$Participant_ID
outs[[8]] <- subset(DATA_wide, Participant_ID %in% Part_out)
names(outs)[8] <- "Valeric_acid"
```



```{r fig.height=15, fig.width=15, warning=FALSE}
for(i in seq(length(outs))){
  sub <- outs[[i]]
  title <- names(outs)[i]
  ll_plot <- list()
  for(i in 1:8){
    ll_plot[[i]] <- plot_outliers_path(sub, DATA_wide, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T)
  }
  
  print(wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") + 
  plot_annotation(title, theme = theme(plot.title = element_text(size = 22, hjust = 0.5))))
}
```


```{r paged.print=FALSE}
outs2 <- sapply(measures, function(i) as.character(unique(outs[[i]]$Participant_ID)))
repet <- table(unlist(outs2))

df <- data.frame(matrix(ncol = length(measures), nrow = length(names(repet))))
rownames(df) <- names(repet)
colnames(df) <- measures

for(i in seq(length(measures))){
  df[outs2[[i]], measures[i]] <- 1
}
df$total <- rowSums(df, na.rm=T)
df[order(df$total, decreasing=T),]
```


## Filtering

```{r}
data_filt <- DATA_wide

id <- which(data_filt$Participant_ID == "ID099")
data_filt$Methylbutyric_acid2[id] <- NA
data_filt$Isobutyric_acid[id] <- NA
data_filt$Isovaleric_acid[id] <- NA

id <- which(data_filt$Participant_ID == "ID077" & data_filt$Timepoint =="T1"|
              data_filt$Participant_ID == "ID119" & data_filt$Timepoint =="T4"|
              data_filt$Participant_ID == "ID017" & data_filt$Timepoint =="T4")
data_filt$Acetic_acid[id] <- NA


id <- which(data_filt$Participant_ID == "ID077" & data_filt$Timepoint =="T1")
data_filt$Butyric_acid[id] <- NA
```

```{r fig.height=14, fig.width=12, warning=FALSE}
ll_plot <- list()
for(i in 1:8){
  ll_plot[[i]] <- plot_sample_measure_CO(data_filt, measures[i], "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T) 
}

wrap_plots(ll_plot, nrow=4)  + plot_layout(guides = "collect") 
```



```{r paged.print=FALSE}
shap <- c()
shap_log <- c()
shap_psd_log <- c()
shap_sqrt <- c()
for(i in 1:8){
  x <- data_filt[,measures[i]]
  y <- replace(x, x == 0, min(x[x > 0])/2)
  
  shap <-  c(shap, shapiro.test(x)$p.value)
  shap_log <-  c(shap_log, shapiro.test(log10(y))$p.value)
  shap_psd_log <-  c(shap_psd_log, shapiro.test(pseudo_log(x, base=10))$p.value)
  shap_sqrt <-  c(shap_sqrt, shapiro.test(sqrt(x))$p.value)
}

data.frame(measures, shap, shap_log, shap_psd_log, shap_sqrt)
```

```{r paged.print=FALSE}
trans_filt <- c("sqrt", "sqrt", "sqrt", "psd_log", "sqrt", "sqrt", "sqrt", "sqrt")
names(trans_filt) <- measures
data.frame(trans_filt)
```



# Correlation between metabolites

```{r fig.height=6, fig.width=7}
library(circlize)
library(ComplexHeatmap)

DATA_wide$sexe_num <- as.numeric(DATA_wide$Geschlecht)
MB_cor <- cor(DATA_wide[, c(6:8, 12:20)], use="complete.obs", method = "spearman")
col_fun_div = colorRamp2(c(-1, 1), hcl_palette = "Blue-Red 3") 
Heatmap(MB_cor, col= col_fun_div, column_title = "Metabolites correlation")
```

```{r fig.height=6, fig.width=7}
MB_cor <- cor(DATA_wide[, 12:19], use="complete.obs", method = "spearman")
col_fun_div = colorRamp2(c(-1, 1), hcl_palette = "Blue-Red 3") 
Heatmap(MB_cor, col= col_fun_div, column_title = "Metabolites correlation")
```

```{r fig.height=6, fig.width=12, message=FALSE}
MB_cor_YB <- cor(subset(DATA_wide, Inter_treat=="Before.Yogurt")[, 12:19], use="complete.obs", method = "spearman") 
MB_cor_YA <- cor(subset(DATA_wide, Inter_treat=="After.Yogurt")[, 12:19], use="complete.obs", method = "spearman") 
ht1 <- Heatmap(MB_cor_YB, col= col_fun_div, column_title = "Before Yogurt")
ht2 <- Heatmap(MB_cor_YA, col= col_fun_div, column_title = "After Yogurt", column_order = column_order(ht1))

ht1 + ht2
```

```{r fig.height=6, fig.width=12, message=FALSE, warning=FALSE}
MB_cor_YOB <- cor(subset(DATA_wide, Inter_treat=="Before.YogurtOatmeal")[, 12:19], use="complete.obs", method = "spearman") 
MB_cor_YOA <- cor(subset(DATA_wide, Inter_treat=="After.YogurtOatmeal")[, 12:19], use="complete.obs", method = "spearman") 
ht3 <- Heatmap(MB_cor_YOB, col= col_fun_div, column_title = "Before Yogurt Oatmeal", column_order = column_order(ht1))
ht4 <- Heatmap(MB_cor_YOA, col= col_fun_div, column_title = "After Yogurt Oatmeal", column_order = column_order(ht1))

ht3 + ht4
```


```{r fig.height=12, fig.width=10, warning=FALSE}
measures_ord <- c("Isobutyric_acid", "Isovaleric_acid", "Methylbutyric_acid2", "Valeric_acid",  "Hexanoic_acid", "Propionic_acid", "Butyric_acid", "Acetic_acid")

pplist <- list()
for(i in 1:8){
  for(j in 1:8){
  z <- i + ((j-1)*8)  
  pplist[[z]] <- plot_pair_measures(DATA_wide, measures_ord[i],  measures_ord[j], subject_name = F, facet_grid = F, scale_log = F)
  }
}

# wrap_plots(pplist,  ncol=8) +  plot_layout(guides = "collect") #saved in Correlations_metabolites.tiff
```

```{r fig.height=12, fig.width=10, warning=FALSE}
pplist <- list()
for(i in 1:8){
  for(j in 1:8){
  z <- i + ((j-1)*8)  
  pplist[[z]] <- plot_pair_measures(DATA_wide, measures_ord[i],  measures_ord[j], subject_name = F, scale_log = F)
  }
}

# wrap_plots(pplist,  ncol=8) #saved in Correlations_metabolites_Inter.tiff
```





# Our mixed model

## 1st version

```{r eval=FALSE, include=FALSE}
# # Check covariates
# lmer1 <- lmer(Valeric_acid ~  Timepoint + Intervention + Alter + Geschlecht + (1|Participant_ID), data=DATA_wide, REML = FALSE)
# lmer2 <- lmer(Valeric_acid ~   Timepoint + Intervention + Alter + (1|Participant_ID), data=DATA_wide, REML = FALSE)
# anova(lmer2, lmer1) # Geschlecht can be remove
# 
# lmer1 <- lmer(Valeric_acid ~  Timepoint + Intervention + Alter  + (1|Participant_ID), data=DATA_wide, REML = FALSE)
# lmer2 <- lmer(Valeric_acid ~   Timepoint + Intervention  + (1|Participant_ID), data=DATA_wide, REML = FALSE)
# anova(lmer2, lmer1) # Alter can be remove
```

```{r}
measures <- colnames(DATA_wide)[12:19]
DF_subs <-  subset_DATA(DATA_wide, measures)
```

```{r}
### Model 1
groupvars <- c("Gruppe", "Timepoint", "Intervention", "(1|Participant_ID)")
res_After <- combined_results(measures, groupvars, DF=DF_subs$DF_A, compar_var = "Intervention",  savingName = "Model1_After", plot_res=T, path=paste(path_analysis, "ResultsMetabolomic/", sep=""))

### Model 2
groupvars <- c("Gruppe", "Timepoint", "Intervention", "baseline_diff", "(1|Participant_ID)")
res_After_full <- combined_results(measures, groupvars, DF=DF_subs$DF_A_full, compar_var = "Intervention", baseline = T,  savingName = "Model2_AfterwBase", plot_res=T, path=paste(path_analysis, "ResultsMetabolomic/", sep=""))

### Model 3-4
groupvars <- c("Gruppe", "Treatment", "(1|Participant_ID)")
res_Yogurt  <- combined_results(measures, groupvars, DF=DF_subs$DF_Y, compar_var = "Treatment",  savingName = "Model3_Yogurt", plot_res=T, path=paste(path_analysis, "ResultsMetabolomic/", sep=""))
res_YogurtOatmeal   <- combined_results(measures, groupvars, DF=DF_subs$DF_YO, compar_var = "Treatment",  savingName = "Model4_YogurtOatmeal", plot_res=T, path=paste(path_analysis, "ResultsMetabolomic/", sep=""))
```



## 2nd version

### no transformation
```{r}
DF_subs <-  prepare_DATA(DATA_wide, NULL)
measures_noTrans <- measures
```


```{r paged.print=FALSE}
groupvars <- c("B", "T", "C1", "C2", "(1|Participant_ID)")

res_After <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF, compar_var = "T", 
                       savingName = "Model_full", plot_res=T, 
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans/", sep=""))

groupvars <- c("B", "P", "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF_int1, compar_var = "B", 
                         savingName = "Model_Yogurt", plot_res=T, 
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans/", sep=""))


res_YogurtOatmeal <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF_int2, compar_var = "B", 
                               savingName = "Model_YogurtOatmeal", plot_res=T, 
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans/", sep=""))
```



```{r paged.print=FALSE}
inf_points_obs <- read.xlsx(file=paste(path_analysis , "ResultsMetabolomic/Rosner_Model_noTrans/Model_full_results.xlsx", sep=""), sheetName="inf_obs_all")
inf_points_obs$NA. <- NULL
# inf_points_obs <- res_After[["inf_obs_all"]]

inf_points_obs <- inf_points_obs[order(inf_points_obs$cooksd, decreasing=T),]
inf_points_obs[1:10,]
```

```{r paged.print=FALSE}
inf_points <- read.xlsx(file=paste(path_analysis , "ResultsMetabolomic/Rosner_Model_noTrans/Model_full_results.xlsx", sep=""), sheetName="inf_part_all")
inf_points$NA. <- NULL
# inf_points <- res_After[["inf_part_all"]]

inf_points <- inf_points[order(inf_points$cooksd, decreasing=T),]
inf_points[1:10,]
```


```{r fig.height=12, fig.width=15, warning=FALSE}
Inf_point <- merge(DATA_wide, inf_points[, c("Participant_ID", "outcome_i")], by=c("Participant_ID"))
sub <- split(Inf_point, Inf_point$outcome_i)
sub <- sub[measures_noTrans] #reorder to correspond to trans
sub[sapply(sub, is.null)] <- NULL # if no outliers detected for one or more measures

ll_plot <- list()
for(i in seq(length(sub))){
  ll_plot[[i]] <- plot_outliers_path(sub[[i]], DATA_wide, names(sub[i]), "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T)
}

wrap_plots(ll_plot, nrow=3)  + plot_layout(guides = "collect") 

```

### no transformation + covariate

3 participants (ID071, ID090, ID113) are removed of the analysis as we do not have their BMI information.

```{r}
### saved in Rosner_Model_noTrans_cov
# same DF and measures_Trans as defined in previous section
```

```{r paged.print=FALSE}

groupvars <- c("B", "T", "C1", "C2", "Alter", "BMI", "Geschlecht", "(1|Participant_ID)")

res_After <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF, compar_var = "T", 
                       savingName = "Model_full", plot_res=T, 
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans_cov/", sep=""))

groupvars <- c("B", "P","Alter", "BMI", "Geschlecht",  "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF_int1, compar_var = "B", 
                         savingName = "Model_Yogurt", plot_res=T, 
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans_cov/", sep=""))


res_YogurtOatmeal <- run_model(measures_noTrans, groupvars, DF=DF_subs$DF_int2, compar_var = "B", 
                               savingName = "Model_YogurtOatmeal", plot_res=T, 
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_noTrans_cov/", sep=""))
```



```{r paged.print=FALSE}
inf_points_obs <- read.xlsx(file=paste(path_analysis , "ResultsMetabolomic/Rosner_Model_noTrans_cov/Model_full_results.xlsx", sep=""), sheetName="inf_obs_all")
inf_points_obs$NA. <- NULL

# inf_points_obs <- res_After[["inf_obs_all"]]
inf_points_obs <- inf_points_obs[order(inf_points_obs$cooksd, decreasing=T),]
inf_points_obs[1:10,]
```

```{r paged.print=FALSE}
inf_points <- read.xlsx(file=paste(path_analysis , "ResultsMetabolomic/Rosner_Model_noTrans_cov/Model_full_results.xlsx", sep=""), sheetName="inf_part_all")
inf_points$NA. <- NULL

# inf_points <- res_After[["inf_part_all"]]
inf_points <- inf_points[order(inf_points$cooksd, decreasing=T),]
inf_points[1:10,]
```




```{r fig.height=12, fig.width=15, warning=FALSE}
Inf_point <- merge(DATA_wide, inf_points[, c("Participant_ID", "outcome_i")], by=c("Participant_ID"))
sub <- split(Inf_point, Inf_point$outcome_i)
sub <- sub[measures_noTrans] #reorder to correspond to trans
sub[sapply(sub, is.null)] <- NULL # if no outliers detected for one or more measures

ll_plot <- list()
for(i in seq(length(sub))){
  ll_plot[[i]] <- plot_outliers_path(sub[[i]], DATA_wide, names(sub[i]), "", x="Timepoint", colorInter = T, print_table=F, trans = trans[i], line_diff = T)
}

wrap_plots(ll_plot, nrow=3)  + plot_layout(guides = "collect") 

```

### transformation

After checking residuals from Rosner_Model_noTrans:

* Acetic_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> overfitting for Alter, negative correlation with Alter and positive correlation with BMI  
* Butyric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter
* Hexanoic_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> bimodality in the data
* Isobutyric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Isovaleric_acid: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Methylbutyric_acid2: 
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, negative correlation with Alter and positive correlation with BMI 
* Propionic_acid: 
  * heteroscedasticity, use box-cox transformation
  * no general good fit for high value, non homogeneity of variance for Geschlecht, cubic relation with Alter and BMI, negative correlation with Alter and positive correlation with BMI 
* Valeric_acid:
  * heteroscedasticity, use box-cox transformation
  * -> non homogeneity of variance for Geschlecht, overfitting ALter, BMI, negative correlation with Alter and positive correlation with BMI 
  
  
  

  
```{r paged.print=FALSE}
### saved in Rosner_Model_Trans_cov

# transformation
trans_analysis <- c("bc", "bc", "bc", "bc", "bc", "bc", "bc", "bc")
names(trans_analysis) <- measures
data.frame(trans_analysis)
```

```{r}
DF_subs <-  prepare_DATA(DATA_wide, trans_analysis)
measures_Trans <- sapply(names(trans_analysis), function(x) if(trans_analysis[[x]] !="") paste(x, trans_analysis[[x]], sep="_") else x)
```

```{r paged.print=FALSE}
groupvars <- c("B", "T", "C1", "C2", "Alter", "BMI", "Geschlecht", "(1|Participant_ID)")

res_After <- run_model(measures_Trans, groupvars, DF=DF_subs$DF, compar_var = "T", 
                       savingName = "Model_full", plot_res=T, 
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/", sep=""))

groupvars <- c("B", "P", "Alter", "BMI", "Geschlecht", "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/", sep=""))


res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/", sep=""))
```



```{r}
### newest model saved in lmer_Trans_cov/
### same results than Rosner_Model_Trans_cov/ but better plots and results presentation

#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2", "Alter", "BMI", "Geschlecht")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P", "Alter", "BMI", "Geschlecht")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P", "Alter", "BMI", "Geschlecht")
RE <- "Participant_ID"
```



```{r}
sig_threshold <- 0.05
wd0 <- paste(path_analysis, "ResultsMetabolomic/", sep="")
setwd(wd0)

run_uniMarker_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work,
               compar_var=compar_var, var_check=NULL, 
               savingName=paste("lmer", compa, sep="_"), 
               sig_threshold=sig_threshold, check_mod=T, save_mod=F,  
               path=paste("lmer_Trans_cov/", compa, "/", sep=""),
               verbose=F)
```



#### Investigate bimodality in Hexanoic_acid
```{r}
DATA_wide_trans <- DF_subs$DF
```

```{r}
p1 <- ggplot(DATA_wide_trans, aes(x=log2(Hexanoic_acid))) + geom_histogram(bins=30, color="white")
p2 <- ggplot(DATA_wide_trans, aes(x=Hexanoic_acid_bc)) + geom_histogram(bins=30, color="white")
wrap_plots(p1, p2, ncol=1)
```


```{r}
DATA_wide_trans$Hexa_group <- ifelse(DATA_wide_trans$Hexanoic_acid_bc >= 5, "1", "0")
```

```{r}
To_export <- DATA_wide_trans[, c("Participant_ID", "Timepoint", "Hexa_group", "Hexanoic_acid_bc")]
write.xlsx(To_export, 
           file="Q:/IPE-A-Epidemiologie/Virginie_Stanislas/Projects/2206_JoghurtStudy/Analysis/DataAnalysis/grouping.xlsx", sheet="Hexanoic")
```



```{r}
ggplot(DATA_wide_trans[-which(is.na(DATA_wide_trans$Hexanoic_acid_bc)),], aes(x=Hexanoic_acid_bc)) + geom_histogram(bins=30, color="white") + facet_wrap(~Hexa_group)
```

```{r fig.height=32, fig.width=12, warning=FALSE}
pp <-list()
for(i in 1:8){
  
  ylim <- range(DATA_wide_trans[, measures[i]], na.rm = T)
  
  ind_filter <- quote(Hexa_group=="0")
  p1 <- plot_sample_measure_CO(DATA_wide_trans, measures[i], "Hexa0", x="Treatment", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim) 
  
  ind_filter <- quote(Hexa_group=="1")
  p2 <- plot_sample_measure_CO(DATA_wide_trans,  measures[i], "Hexa1", x="Treatment", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim)
  
  
  pp[[i]] <- wrap_plots(p1, p2, ncol=2)
  
}

wrap_plots(pp, ncol=1)
```


```{r fig.height=32, fig.width=12, warning=FALSE}
pp <-list()
for(i in 1:8){
  
  ylim <- range(DATA_wide_trans[, measures[i]], na.rm = T)
  
  ind_filter <- quote(Hexa_group=="0")
  p1 <- plot_sample_measure_CO(DATA_wide_trans, measures[i], "Hexa0", x="Timepoint", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim) 
  
  ind_filter <- quote(Hexa_group=="1")
  p2 <- plot_sample_measure_CO(DATA_wide_trans,  measures[i], "Hexa1", x="Timepoint", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim)
  
  
  pp[[i]] <- wrap_plots(p1, p2, ncol=2)
  
}

wrap_plots(pp, ncol=1)
```





```{r}
var <- colnames(DATA_wide_trans)[c(4:8,26:34)]
tbl_summary(DATA_wide_trans[, var], by=Hexa_group, missing ="no") %>% add_p() %>% 
  gtsummary::as_gt()
```


```{r}
DATA_wide_trans$Hexanoic_acid_bc <- NULL
DATA_wide_trans$Methylbutyric_acid2_bc <- NULL
DATA_wide_trans$Acetic_acid_bc <- NULL
DATA_wide_trans$Butyric_acid_bc <- NULL
DATA_wide_trans$Isobutyric_acid_bc <- NULL
DATA_wide_trans$Isovaleric_acid_bc <- NULL
DATA_wide_trans$Propionic_acid_bc <- NULL
DATA_wide_trans$Valeric_acid_bc <- NULL

Hexa_sub1 <- subset(DATA_wide_trans, Hexa_group=="1")
Hexa_sub0 <- subset(DATA_wide_trans, Hexa_group=="0")

# trans_hexa <- c("bc") # we only transform Hexanoic_acid but keep the previous transformation stored in DATA_wide_trans for the other SCFAs 
# names(trans_hexa) <- "Hexanoic_acid" 
DF_subs_Hexa1 <- prepare_DATA(Hexa_sub1, trans_analysis)
DF_subs_Hexa0 <- prepare_DATA(Hexa_sub0, trans_analysis)
```



```{r paged.print=FALSE}
groupvars <- c("B", "T", "C1", "C2", "Alter", "BMI", "Geschlecht", "(1|Participant_ID)")

res_After <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF, compar_var = "T",
                       savingName = "Model_full_Hexa1", plot_res=T,
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))


res_After <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF, compar_var = "T",
                       savingName = "Model_full_Hexa0", plot_res=T,
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))


groupvars <- c("B", "P", "Alter", "BMI", "Geschlecht", "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt_Hexa1", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))

res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt_Hexa0", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))

res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal_Hexa1", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))

res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal_Hexa0", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans_cov/Hexa_sub/", sep=""))
```




### transformation no covariates

```{r}
### saved in Rosner_Model_Trans
# same DF and measures_Trans as defined in previous section (before section on Hexanoic_acid bimodality)
```

```{r paged.print=FALSE}

groupvars <- c("B", "T", "C1", "C2", "(1|Participant_ID)")

res_After <- run_model(measures_Trans, groupvars, DF=DF_subs$DF, compar_var = "T", 
                       savingName = "Model_full", plot_res=T, 
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/", sep=""))

groupvars <- c("B", "P", "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/", sep=""))


res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/", sep=""))
```


```{r}
### newest model saved in lmer_Trans/
### same results than Rosner_Model_Trans/ but better plots and results presentation

#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
```

```{r}
sig_threshold <- 0.05
wd0 <- paste(path_analysis, "ResultsMetabolomic/", sep="")
setwd(wd0)

run_uniMarker_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work,
               compar_var=compar_var, var_check=NULL, 
               savingName=paste("lmer", compa, sep="_"), 
               sig_threshold=sig_threshold, check_mod=T, save_mod=F,  
               path=paste("lmer_Trans/", compa, "/", sep=""),
               verbose=F)
```



#### bimodality in Hexanoic_acid

```{r}
DATA_wide_trans$Hexanoic_acid_bc <- NULL
DATA_wide_trans$Methylbutyric_acid2_bc <- NULL
DATA_wide_trans$Acetic_acid_bc <- NULL
DATA_wide_trans$Butyric_acid_bc <- NULL
DATA_wide_trans$Isobutyric_acid_bc <- NULL
DATA_wide_trans$Isovaleric_acid_bc <- NULL
DATA_wide_trans$Propionic_acid_bc <- NULL
DATA_wide_trans$Valeric_acid_bc <- NULL

Hexa_sub1 <- subset(DATA_wide_trans, Hexa_group=="1")
Hexa_sub0 <- subset(DATA_wide_trans, Hexa_group=="0")

# trans_hexa <- c("bc") # we only transform Hexanoic_acid but keep the previous transformation stored in DATA_wide_trans for the other SCFAs 
# names(trans_hexa) <- "Hexanoic_acid" 
DF_subs_Hexa1 <- prepare_DATA(Hexa_sub1, trans_analysis)
DF_subs_Hexa0 <- prepare_DATA(Hexa_sub0, trans_analysis)
```



```{r paged.print=FALSE}
groupvars <- c("B", "T", "C1", "C2", "(1|Participant_ID)")

res_After <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF, compar_var = "T",
                       savingName = "Model_full_Hexa1", plot_res=T,
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))


res_After <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF, compar_var = "T",
                       savingName = "Model_full_Hexa0", plot_res=T,
                       path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))


groupvars <- c("B", "P", "(1|Participant_ID)")
res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt_Hexa1", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))

res_Yogurt  <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF_int1, compar_var = "B",
                         savingName = "Model_Yogurt_Hexa0", plot_res=T,
                         path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))

res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa1$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal_Hexa1", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))

res_YogurtOatmeal <- run_model(measures_Trans, groupvars, DF=DF_subs_Hexa0$DF_int2, compar_var = "B",
                               savingName = "Model_YogurtOatmeal_Hexa0", plot_res=T,
                               path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_Trans/Hexa_sub/", sep=""))
```



### Multiple outliers filtered

```{r}
# WE DID NOT KEEP THIS ANALYSIS

# DF_filt_subs <-  prepare_DATA(data_filt, trans_filt)
# measures2 <- sapply(names(trans_filt), function(x) if(trans_filt[[x]] !="") paste(x, trans_filt[[x]], sep="_") else x)
```

```{r}
# groupvars <- c("B", "T", "C1", "C2", "(1|Participant_ID)")
# 
# res_After <- run_model(measures2, groupvars, DF=DF_filt_subs$DF, compar_var = "T", 
#                        savingName = "Model_full", plot_res=T, 
#                        path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_filt/", sep=""))
# 
# groupvars <- c("B", "P", "(1|Participant_ID)")
# res_Yogurt  <- run_model(measures2, groupvars, DF=DF_filt_subs$DF_int1, compar_var = "B", 
#                          savingName = "Model_Yogurt", plot_res=T, 
#                          path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_filt/", sep=""))
# 
# 
# res_YogurtOatmeal <- run_model(measures2, groupvars, DF=DF_filt_subs$DF_int2, compar_var = "B", 
#                                savingName = "Model_YogurtOatmeal", plot_res=T, 
#                                path=paste(path_analysis, "ResultsMetabolomic/Rosner_Model_filt/", sep=""))
```






# Participants differences in response


```{r fig.height=32, fig.width=12, warning=FALSE}
## Separate participant increasing during washout from people decreasing during washout
pp <-list()
for(i in 1:8){
  ind_filter <- quote(Timepoint=="T2"& Diff=="Increase")
  
  ylim <- range(DATA_wide[, measures[i]], na.rm = T)
  
  p1 <- plot_sample_measure_CO(DATA_wide, measures[i], "", x="Timepoint", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim) 
  
  ind_filter <- quote(Timepoint=="T2"& Diff=="Decrease")
  p2 <- plot_sample_measure_CO(DATA_wide,  measures[i], "", x="Timepoint", colorInter = T, print_table=F,  trans = trans[i], ind_filter=ind_filter, ylim=ylim)
  
  
  pp[[i]] <- wrap_plots(p1, p2, ncol=2)
  
}

wrap_plots(pp, ncol=1)
```




```{r message=FALSE}
myCol <- brewer.pal(8, "Pastel2")
names(myCol) <- measures
DF <- DATA_wide

ind_filter <- list()
for(i in 1:8){
  DF_F <- computeDiff(DF, "Timepoint", measures[i])
  ind_filter[[i]] <- as.character(subset(DF_F, Timepoint=="T2"& Diff=="Increase")$Participant_ID)
}
names(ind_filter) <- measures

plt <- venn.diagram(x = ind_filter[1:5],
                    disable.logging=T,
                    category.names = measures[1:5],
                    lty = 'blank', fill = myCol[1:5], # circle
                    fontface = "bold", fontfamily = "sans", # Numbers
                    cat.fontface = "bold", cat.fontfamily = "sans", # Set names 
                    margin =0.3,filename=NULL)

plt2 <- venn.diagram(x = ind_filter[c(6,7,3,4,8)], 
                     disable.logging=T,
                    category.names = measures[c(6,7,3,4,8)],
                    lty = 'blank', fill = myCol[c(6,7,3,4,8)], # circle
                    fontface = "bold", fontfamily = "sans", # Numbers
                    cat.fontface = "bold", cat.fontfamily = "sans", # Set names
                    margin =0.3, filename=NULL)


```

```{r fig.height=6, fig.width=7}
grid.newpage()
grid::grid.draw(plt)
grid.newpage()
grid::grid.draw(plt2)
```


```{r}
length(Reduce(intersect,ind_filter)) # 14 # number of individuals find increasing during wahsout for all metabolite
```




```{r}
# Ven diagram depending on correlation
plt <- venn.diagram(x = ind_filter[c("Isobutyric_acid", "Isovaleric_acid", "Methylbutyric_acid2")], 
                    disable.logging=T,
                    category.names = c("Isobutyric_acid", "Isovaleric_acid", "Methylbutyric_acid2"),
                    lty = 'blank', 
                    fill = myCol[c("Isobutyric_acid", "Isovaleric_acid", "Methylbutyric_acid2")], # circle
                    fontface = "bold", fontfamily = "sans", # Numbers
                    cat.fontface = "bold", cat.fontfamily = "sans", # Set names
                    margin =0.3, filename=NULL)

plt2 <- venn.diagram(x = ind_filter[c("Propionic_acid", "Butyric_acid", "Acetic_acid")], 
                    disable.logging=T,
                    category.names = c("Propionic_acid", "Butyric_acid", "Acetic_acid"),
                    lty = 'blank', 
                    fill = myCol[c("Propionic_acid", "Butyric_acid", "Acetic_acid")], # circle
                    fontface = "bold", fontfamily = "sans", # Numbers
                    cat.fontface = "bold", cat.fontfamily = "sans", # Set names
                    margin =0.3, filename=NULL)

# 
plt3 <- venn.diagram(x = ind_filter[c("Isobutyric_acid", "Butyric_acid", "Hexanoic_acid")], 
                    disable.logging=T,
                    category.names = c("Isobutyric_acid", "Butyric_acid", "Hexanoic_acid"),
                    lty = 'blank', 
                    fill = myCol[c("Isobutyric_acid", "Butyric_acid", "Hexanoic_acid")], # circle
                    fontface = "bold", fontfamily = "sans", # Numbers
                    cat.fontface = "bold", cat.fontfamily = "sans", # Set names
                    margin =0.3, filename=NULL)
```


```{r fig.height=4, fig.width=5}
grid.newpage()
grid::grid.draw(plt)

grid.newpage()
grid::grid.draw(plt2)

grid.newpage()
grid::grid.draw(plt3)
```


