---
title: "Taxonomic Analysis of relative values with all samples"
output: 
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---


<style>
.main-container { width: 1200px; max-width:2800px;}
</style>




# Loading

```{r message=FALSE}
library(phyloseq)
library(ggplot2)
library(plyr)
library(ape)
library(knitr)
library(vegan)
library(tidyr)
library(gridExtra)
# library(lemon)
library(RColorBrewer)
library(dplyr)
library(cowplot)
library(scater)
library(ggnewscale)
library(mia)
library(miaViz)
library(kableExtra)
library(patchwork)
library(microViz)
library(lme4)
library(emmeans)
library(flextable)
library(ggpubr)
library(reshape)
library(tictoc)
```


```{r}
dir <- "GitHub_repository/IPE_yogurt_rolledOat_microbiome/"
path_analysis <- paste(dir, "DataAnalysis/Results_Main_Analysis/ResultsTaxonomy/", sep="")

 

source(file = paste(dir, "R functions/Utils.R", sep=""))
source(file = paste(dir, "R functions/Utils_metagenomics.R", sep=""))
source(file = paste(dir, "R functions/Utils_longitudinal.R", sep=""))
source(file = paste(dir, "R functions/Utils_crossOver.R", sep=""))
source(file = paste(dir, "R functions/code_utils.R", sep=""))
```




```{r}
### Load data with already ordination and alpha diversity (saved at step 2.3.1)
### contains tse, tse_sub, tse_SGB
load(paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
```

# Data preparation

```{r message=FALSE}
load(paste(dir, "DataPreparation/Microbiome/Taxonomy/Data_cleaned_ALL_unfiltered.Rdata", sep=""))
load(paste(dir, "DataPreparation/Microbiome/Taxonomy/Data_cleaned_ALL_filtered.Rdata", sep=""))

# Define working objects
tse_SGB0 <- tse
physeq_SGB0 <- physeq_rel_SGBR

tse0 <- altExp(tse,"Species")  # 1366 440 
physeq0 <- physeq_rel_spR


tse_SGB_sub0 <- tse_clean

tse_sub0 <- altExp(tse_clean,"Species") # 698 440  
```

```{r}
# Remove 2 samples from ID016 not correctly prepared by the participant
tse_SGB <- tse_SGB0[, c(!(tse_SGB0$Sample_ID  %in% c("JH379", "JH380")))]
tse <- tse0[, c(!(tse0$Sample_ID  %in% c("JH379", "JH380")))]

tse_SGB_sub <- tse_SGB_sub0[, c(!(tse_SGB_sub0$Sample_ID  %in% c("JH379", "JH380")))]
tse_sub <- tse_sub0[, c(!(tse_sub0$Sample_ID  %in% c("JH379", "JH380")))]

physeq_SGB <- ps_filter(physeq_SGB0, !(Sample_ID %in% c("JH379", "JH380")))
physeq <- ps_filter(physeq0, !(Sample_ID %in% c("JH379", "JH380")))
```


```{r message=FALSE}
# removed ordered factor to allow some plotting functions to work properly
physeq_SGB_noFact <-  physeq_SGB
sample_data(physeq_SGB_noFact)$Inter_treat <- as.character(sample_data(physeq_SGB_noFact)$Inter_treat )
sample_data(physeq_SGB_noFact)$Inter_treat <- as.factor(sample_data(physeq_SGB_noFact)$Inter_treat )
```


```{r}
cols <- c("#017b74", "#73ffec", "#ad0053", "#ffcacf", "#ff8d3f", "#ffdf91", "#00c14e","#abffc2", "#6042ed", "#90b9ff", "#700054","#ff8bfc", "#3d8b00", "#bcff7c", "#8b3700", "#ffba91", "#ff4e2a", "#000e26", "#71a800", "#577eff", "#02d0d3", "#d00039", "#dba500", "#ac72ff", "#e0c9ff", "#002e66", "#006dbd", "#02a0c8", "#db37ec",  "#e5ffd5", "#004a20", "#637500", "#ec00bf", "#434d00","#ff77a8", "#002f17", "#f1ff4b", "#533500", "#7ea2ff", "#000d4e", "#ff41a0", "#003636", "#271800", "#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#A65628", "#F781BF", "#FFFF00FF", "#999999")

cols_Inter_treat <- c(After.Yogurt="#7570B3", After.YogurtOatmeal="#D95F02", Before.Yogurt="#AAA8C5", Before.YogurtOatmeal="#DBB394")

cols_Participant_Id <- rep(cols, 3)
names(cols_Participant_Id) <- unique(colData(tse)[, "Participant_ID"])
```

# Exploration

## Prevalence x Abundance

**prevalence**: frequency of samples where the taxa was detected

**abundance**: number of reads assigned to a taxa (on average or in total)

```{r}
rowData(tse)$prevalence <- getPrevalence(tse, detection = 0, sort = FALSE, assay.type = "relabundance", as_relative = TRUE)
rowData(tse)$abundance_sum <- rowSums(assays(tse)[["relabundance"]])
rowData(tse)$abundance_mean <- rowMeans(assays(tse)[["relabundance"]])
rowData(tse)$abundance_max <- apply(assays(tse)[["relabundance"]], 1, max)
```



```{r fig.height=10, fig.width=12}
## Prevalence x abundance on the complete species data - axis log transformed
ggplot(as.data.frame(rowData(tse)), aes(abundance_sum, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), labels = scales::label_number(), name = "Total Abundance") +
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * ncol(tse), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )) + 
  ggrepel::geom_text_repel(
    #data = function(df) filter(df, abundance_sum > 300 | prevalence > 0.6),
  mapping = aes(label = Species), size = 2.5, min.segment.length = 0)+
   theme_bw() 
```


# Ordination



- **Distance/dissimilarity measures** 
  - **`r colorize("Bray-curtis", "#02a0c8")`**: take into account the abundance $BC_{ij}= 1 - \frac{2C_{ij}}{S_i + S_j}$ with $S_i$ : Total number of species in sample $i$, $S_j$ : Total number of species in sample $j$ and $C_{ij}$ : Sum of lesser value for those species common in both samples
  - **`r colorize("Unifrac", "#02a0c8")`**:  take into account the phylogenetic tree information
      - unweigthed unifrac: Measures the fraction of unique branch length / measures the relative length of those branches that lead exclusively to species present in only one of the two samples with respect to the total length of all branches in the tree.
    - weigthed unifrac: includes information on the relative abundance of each taxa
  - **`r colorize("Jaccard", "#02a0c8")`**: presence / absence, ignores abundance and phylogenetic information
  - **`r colorize("Aitchison", "#02a0c8")`**: take into consideration the compositional structure of the data. Dissimilarity measure calculated as the Euclidean distance between samples after performing a centered log ratio (clr) transformation.
  
- **Ordination/dimention reduction methods** 
  - **`r colorize("PCoA/MDS", "#02a0c8")`**: extension of Principal Components Analysis (PCA),  detect linear trends
  - **`r colorize("NMDS", "#02a0c8")`**: mon-metric multidimensional scaling can detect non-linear patterns
  - **`r colorize("TSNE", "#02a0c8")`**
  - **`r colorize("UMAP", "#02a0c8")`**
  


```{r}
load(paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
```


```{r}
tse <- runMDS(tse, FUN = vegan::vegdist, method = "bray", name = "PCoA_BC", exprs_values = "relabundance", keep_dist = T)
set.seed(40)
tse <- runMDS(tse, FUN = mia::calculateUnifrac,  weighted=T, name = "PCoA_wUnifrac", tree = rowTree(tse), ntop = nrow(tse), exprs_values = "relabundance", keep_dist = T)

tse <- runNMDS(tse, FUN = vegan::vegdist, method = "bray", name = "NMDS_BC", exprs_values = "relabundance", keep_dist = T)
set.seed(40)
tse <- runNMDS(tse, FUN = mia::calculateUnifrac, weighted=T, name = "NMDS_wUnifrac", tree = rowTree(tse), ntop = nrow(tse), exprs_values = "relabundance", keep_dist = T)


set.seed(122)
tse <- runTSNE(tse, name = "TSNE", exprs_values="relabundance")
set.seed(122)
tse <- runUMAP(tse, name = "UMAP", exprs_values="relabundance")
```



## UMAP

```{r fig.height=12, fig.width=24, message=FALSE, warning=FALSE}
pp <- plot_ordination(tse, "UMAP", sd_threshold=1.4, rank_dominance="Family")
PartHsd_UMAP <- pp$Participants_high_sd
p1 <- pp$p_gruppe #850*550
p2 <- pp$p_topdom
p3 <- plot_ordination(tse, "UMAP", sd_threshold=1.4, rank_dominance="Genus")$p_topdom #1100*550
p4 <- pp$p_inter_treat #900*550
p5 <- pp$p_part #900*550
p6 <- pp$p_hv_part

#(p1 + p5 +  p2) / (p4 + p6 + p3) + plot_layout(guides = 'collect')
```

```{r fig.height=9, fig.width=14, message=FALSE, paged.print=FALSE}
p3
```

```{r fig.height=24, fig.width=24, message=FALSE, warning=FALSE, paged.print=FALSE}
genus_multi <- c("Prevotella", "Bacteroides", "Ruminococcus", "Bifidobacterium","Clostridia_unclassified", "Phocaeicola", "Roseburia", "Candidatus_Cibiobacter", "Clostridium", "Coprococcus", "Faecalibacterium", "Lachnospiraceae_unclassified")

p_list <- list()
for(i in seq(length(genus_multi))){
genus <- genus_multi[i]
p_list[[i]] <- plot_ordination(tse, "UMAP", genus=genus)$p_1spec 
}
#wrap_plots(p_list, ncol=3) 
```

  
# Alpha diversity on Species space, unfiltered


- **Richness** refers to the number of species in a sample:
  - **`r colorize("Observed richness", "#02a0c8")`**: number of observed species $S$
    - the most sensitive to the difference in sampling effort, since it weights all species equally independent from their relative abundances
  - **`r colorize("Chao1", "#02a0c8")`** assumes that rare species carry information about the (unknown) number of unobserved species. 
    - $S_{Chao1} = S + \frac{n_1(n_1 - 1)}{2(n_2 + 1)}$ with $S$ the total number of observed species, $n_1$ singletons and $n_2$ doubletons
    - Cannot be computed on relative abundance, needs counts values
- **Evenness** refers to the distribution of the species in the sample: how similar the abundances of the different species are.
  - **`r colorize("Pielou", "#02a0c8")`** corresponds to the Shannon diversity index normalized by the observed richness: $H/ln(S)$
  - **`r colorize("Simpson evenness", "#02a0c8")`** corresponds to the inverse Simpson diversity normalized by the observed richness $Sim/S$.
- **Diversity** indices measure the overall community heterogeneity depending on both species richness and evenness.
  - **`r colorize("Hill coefficient", "#02a0c8")`** combines many standard indices into a single equation 
    - indicates the number of species whose even distribution would lead to the same diversity than the observed community, where the species abundances are unevenly distributed.
  - **`r colorize("Shannon", "#02a0c8")`**  
    - $H=-\sum^S_{i=1}p_iln(p_i)$ with $S$ total number of species, $p_i$ proportion of species $i$
  - **`r colorize("Inverse Simpson", "#02a0c8")`**  
    - $Sim=1/\sum p_i^2$ with $p_i$ proportion of species $i$  
- **Phylogenetic diversity** incorporates information from phylogenetic relationships between species in the sample. 
  - **`r colorize("Faith's Phylogenetic diversity", "#02a0c8")`** measures how long the taxonomic distance is between taxa that are present in the sample. 
    - sum of branch length of all species in a sample. Larger values represent higher diversity.
- **Dominance** are in general negatively correlated with diversity. High dominance is obtained when one or few species have a high share of the total species abundance in the sample.
  - **`r colorize("dbp", "#02a0c8")`**: relative abundance of the most abundant species of the sample 
    - $R=N_1/N_{tot}$, with $N_1$ the absolute abundance of the most dominant species and $N_{tot}$ the sum of absolute abundances of all species in the sample. 
    - Note: here as the value are already relative, this measure correspond simply to the absolute index $N_1$ (abundance of the most dominant species) / 100
  - **`r colorize("Core abundance", "#02a0c8")`**: relative abundance of core species species in the sample  
    - $N_{core}/N_{tot}$, with $N_{core}$ the absolute abundance of core species and $N_{tot}$ the sum of absolute abundances of all species in the sample. 
    - Core species are defined as those species that have prevalence over 50 in the dataset.
- **Rarity** indices characterize the concentration of taxa at low abundance.
  - **`r colorize("Log modulo skewness", "#02a0c8")`** quantifies the concentration of the least abundant species by the log-modulo skewness of the arithmetic abundance classes
- **Divergence** quantifies heterogeneity within a sample by the average sample dissimilarity with a given reference sample.
  - compute distance of each samples against one reference (here median)



```{r}
tse <- estimateRichness(tse, assay_name = "relabundance", index = c( "hill", "observed"), name=c( "Hill", "Observed"))
tse <- estimateEvenness(tse, assay_name = "relabundance", index=c( "pielou", "simpson_evenness"), name =   c("Pielou", "Simpson_evenness"))
tse <- estimateDiversity(tse, assay_name = "relabundance", index = c("shannon", "inverse_simpson") , name = c("Shannon", "Inverse_Simpson"))
tse <- estimateFaith(tse, assay_name = "relabundance", tree=rowTree(tse), name = "Faith")
tse <- estimateDominance(tse, assay_name = "relabundance", index=c("dbp", "core_abundance"), name = c("dbp", "Core_abundance"))
tse <- estimateDiversity(tse, assay_name = "relabundance", index = "log_modulo_skewness", name = "Log_modulo_skewness") # rarity
tse <- estimateDivergence(tse, assay_name = "relabundance", reference = "median", FUN = vegan::vegdist, name="Divergence_to_median") 
```

## Saving tse
```{r}
save(tse, tse_sub, tse_SGB, file=paste(path_analysis, "tse_allwithordination.Rdata", sep=""))
# data with ordination and alpha diversity 
# To keep same ordination (ran on Computer office)
```

## Plots


```{r}
measures <- c("Observed", "Pielou", "Simpson_evenness", "Shannon", "Inverse_Simpson", "Faith", "dbp", "Core_abundance", "Log_modulo_skewness", "Divergence_to_median")
subtitle <- c("Measure of richness", "Measure of Evenness", "Measure of Evenness", "Measure of diversity", "Measure of diversity", "Measure of Phylogenetic diversity", "Measure of Dominance", "Measure of Dominance", "Measure of rarity", "Measure of divergence")

# By intervention
pp_list <- list()
for(i in seq(length(measures))){
  pp_list[[i]] <- plot_sample_measure_CO(tse, measures[i], subtitle[i])
}
```

```{r fig.height=12, fig.width=14}
wrap_plots(pp_list[1:6], nrow=3) 
```


```{r fig.height=8, fig.width=14}
wrap_plots(pp_list[7:10], nrow=2) 
```


```{r}
measures <- c("Observed", "Pielou", "Simpson_evenness", "Shannon", "Inverse_Simpson", "Faith", "dbp", "Core_abundance", "Log_modulo_skewness", "Divergence_to_median")
subtitle <- c("Measure of richness", "Measure of Evenness", "Measure of Evenness", "Measure of diversity", "Measure of diversity", "Measure of Phylogenetic diversity", "Measure of Dominance", "Measure of Dominance", "Measure of rarity", "Measure of divergence")

# By time points
pp_list <- list()
for(i in seq(length(measures))){
  pp_list[[i]] <- plot_sample_measure_CO(tse, measures[i], subtitle[i], x="Timepoint", colorInter = T, print_table=F)
}
```

```{r fig.height=12, fig.width=12}
wrap_plots(pp_list[1:6], nrow=3) 
```

```{r fig.height=8, fig.width=12}
wrap_plots(pp_list[7:10], nrow=2) 
```


## Models

After checking residuals:

* Core_abundance:
  * heteroscedasticity toward small values bc
* dbp:
  * heteroscedasticity bc
* Divergence_to_median 
  * skewed to the right
* Faith
  * ok
* Inverse_Simpson
  * small heteroscdasticity
* Log_modulo_skewness
   * heteroscedasticity toward small values bc
* Observed
  * small heteroscdasticity
* Pielou
  *  skewed to the left
* Shannon
  *  skewed to the left
* Simpson_evenness
  * ok


```{r paged.print=FALSE}
### saved in Rosner_Model_Trans_cov
DF <- as.data.frame(colData(tse))

# transformation
trans_analysis <- c("bc", "bc", "", "bc", "bc", "", "bc", "bc", "bc", "bc")
names(trans_analysis) <- measures
data.frame(trans_analysis)
```


```{r}
DF_subs <- prepare_DATA(DF, trans_analysis)
measures_Trans <- sapply(names(trans_analysis), function(x) if(trans_analysis[[x]] !="") paste(x, trans_analysis[[x]], sep="_") else x)
```



```{r}
### newest model saved in lmer_Trans/
### same results than Rosner_Model_Trans/ but better plots and results presentation

#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"
```



```{r}
sig_threshold <- 0.05
wd0 <- paste(path_analysis, "ResultsAlpha/", sep="")
setwd(wd0)

run_uniMarker_model(measures_Trans, package="lme4", counts=F, FE=FE, RE=RE, DF=DF_work,
               compar_var=compar_var, var_check=NULL, 
               savingName=paste("lmer", compa, sep="_"), 
               sig_threshold=sig_threshold, check_mod=T, save_mod=F,  
               path=paste("lmer_Trans/", compa, "/", sep=""),
               verbose=F)
```



# Bar plots on Species space, unfiltered 

```{r fig.height=12, fig.width=12, message=FALSE, warning=FALSE}
n_taxa <- 50
cols_mod <- cols
cols_mod[n_taxa+1] <- "lightgrey"
  

physeq_SGB_noFact  %>%
  ps_select(Participant_ID, Inter_treat) %>% 
  phyloseq::merge_samples(group = "Inter_treat") %>%
  comp_barplot(tax_level = "Species", n_taxa = n_taxa, bar_width = 0.9, palette=cols_mod, 
               tax_transform_for_plot="identity", bar_outline_colour = NA, 
               sample_order=c("Before.Yogurt", "After.Yogurt", "Before.YogurtOatmeal", "After.YogurtOatmeal")) + 
  labs(x = NULL, y = NULL) + guides(fill = guide_legend(ncol = 1, reverse=TRUE)) +theme(legend.key.size = unit(0.5, 'cm'))


physeq_SGB_noFact  %>%
  ps_select(Participant_ID, Inter_treat) %>% 
  phyloseq::merge_samples(group = "Inter_treat") %>%
  comp_barplot(tax_level = "Genus", n_taxa = n_taxa, bar_width = 0.9, palette=cols_mod, 
               tax_transform_for_plot="identity", bar_outline_colour = NA, 
               sample_order=c("Before.Yogurt", "After.Yogurt", "Before.YogurtOatmeal", "After.YogurtOatmeal")) + 
  labs(x = NULL, y = NULL) + guides(fill = guide_legend(ncol = 1, reverse=TRUE)) +theme(legend.key.size = unit(0.5, 'cm'))

```


```{r fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
physeq_clust <- physeq %>%
  ps_filter(Timepoint  == "T1") %>% ps_seriate()
Parti_ord <- as.character(sample_data(physeq_clust)$Participant_ID)
sa <- sample_data(physeq)[, c("Participant_ID", "Sample_ID")]
Sample_ord <- sa[order(factor(sa$Participant_ID, levels=Parti_ord)), ]$Sample_ID

physeq_SGB %>%
  comp_barplot("SGB", facet_by  = "Inter_treat", n_taxa = n_taxa, sample_order =Sample_ord,  label = "Participant_ID", bar_outline_colour=NA, bar_width=0.9, tax_transform_for_plot="identity",  palette=cols_mod, ncol=4)+
  coord_flip() + guides(fill = guide_legend(ncol = 1))

physeq %>%
  comp_barplot("Species", facet_by  = "Inter_treat", n_taxa = n_taxa, sample_order =Sample_ord,  label = "Participant_ID", bar_outline_colour=NA, bar_width=0.9, tax_transform_for_plot="identity",  palette=cols_mod, ncol=4)+
  coord_flip() + guides(fill = guide_legend(ncol = 1))

physeq %>%
  comp_barplot("Genus", facet_by  = "Inter_treat", n_taxa = n_taxa, sample_order =Sample_ord,  label = "Participant_ID", bar_outline_colour=NA, bar_width=0.9, tax_transform_for_plot="identity",  palette=cols_mod, ncol=4)+
  coord_flip() + guides(fill = guide_legend(ncol = 1))

physeq %>%
  comp_barplot("Family", facet_by  = "Inter_treat", n_taxa = n_taxa, sample_order =Sample_ord,  label = "Participant_ID", bar_outline_colour=NA, bar_width=0.9, tax_transform_for_plot="identity",  palette=cols_mod, ncol=4)+
  coord_flip() + guides(fill = guide_legend(ncol = 1))
```

```{r fig.height=8, fig.width=15, message=FALSE}
taxa <- c("Streptococcus_thermophilus", "Lactobacillus_delbrueckii")

pp_list <- list()
pp_list2 <- list()
for(i in seq(length(taxa))){
  
  colData(tse)$abundance <- getAbundanceFeature(tse, feature_id = paste("Species:", taxa[i],sep=""), assay.type = "relabundance")
  pp_list[[i]] <- plot_sample_measure_CO(tse, "abundance", taxa[i],  x="Timepoint", colorInter = T, trans = "psdlog", print_table = F)  
  pp_list2[[i]] <- plot_sample_measure_CO(tse, "abundance", taxa[i],  trans = "psdlog")  
}


wrap_plots(c(pp_list, pp_list2), nrow=2) 
```







# Beta diversity

```{r}
library(NBZIMM)
```


## Models

```{r message=FALSE}
# tse_sub2 <- subsetByPrevalentTaxa(tse_sub, detection = 0, prevalence = 0.2, assay_name="relabundance") # 335 438 
tse_sub2 <- tse_sub
assay_t <- t(assays(tse_sub2)[["relabundance"]]) %>% as.data.frame()
colnames(assay_t) <- gsub("Species:", "", colnames(assay_t))
taxa <- colnames(assay_t)
assay_t$Sample_ID <- rownames(assay_t)
metadata <- as.data.frame(colData(tse_sub2))
DF <- merge(metadata, assay_t, by="Sample_ID")
DF_subs <- prepare_DATA(DF, NULL)
```



```{r}
#### Model full
compar_var = "T"
DF_work=DF_subs$DF
compa = "full"
FE <- c("B", "T", "C1", "C2")
RE <- "Participant_ID"

#### Model Y
compar_var = "B"
DF_work=DF_subs$DF_int1
compa = "Y"
FE <- c("B", "P")
RE <- "Participant_ID"

#### Model YO
compar_var = "B"
DF_work=DF_subs$DF_int2
compa = "YO"
FE <- c("B", "P")
RE <- "Participant_ID"

```


```{r}
model <- "NBZIMM_rel_asinsqrt"
trans <- ""
sig_threshold <- 0.01
```

```{r}
DF_work_mod=DF_work
DF_work_mod[,taxa] <- asin(sqrt(DF_work_mod[, taxa]/100))
```


```{r}
tic()
# ZIG AST
setwd(paste(path_analysis, "ResultsBetadiv/", sep=""))
run_taxa_model(taxa, package="NBZIMM", counts=F, FE=FE, RE=RE, DF=DF_work_mod, compar_var=compar_var, var_check=NULL, savingName=paste("NBZIMM_rel_asinsqrt_", compa, sep=""), sig_threshold=sig_threshold, check_mod=F, save_mod=F,  path=paste("NBZIMM_rel_asinsqrt/", compa, "/", sep=""))
toc()
# Model full: 596.56 sec elapsed 240.58 sec elapsed
# Model Y: 417.39 sec elapsed 196.39 sec elapsed
# Model YO: 430.28 sec elapsed 236.4 sec elapsed
```


```{r}
# Convert arcsine-square root–transformed model coefficients for Streptococcus and Lactobacillus

# Coefficient values are extracted from the Excel tables: NBZIMM_rel_asinsqrt_Y_results_prev0.2.xlsx and NBZIMM_rel_asinsqrt_YO_results_prev0.2.xlsx tab= PRS_all

# Back-transformation to relative abundance when baseline =0: 
# RA_pct = (sin(coef))^2 * 100.

# Interpretation:
# - RA_pct is the predicted relative abundance of the taxon expressed as a percentage of the
#   total microbial community in the sample.
# - If baseline abundance = 0, RA_pct gives the exact post-intervention abundance
#   and also corresponds to the absolute change from baseline.
# - If baseline > 0 or unknown, post-intervention abundance should be computed as:
#   RA_post = (sin(y0_trans + coef))^2 * 100
#   Absolute change from baseline:
#   Delta_RA = (sin(y0_trans + coef))^2 - (sin(y0_trans))^2
#   where y0_trans = arcsin(sqrt(baseline))

strep_coef_Y <- 0.0984
strep_coef_Y_low <- 0.0842
strep_coef_Y_high <- 0.1125


strep_coef_YO <- 0.0889
strep_coef_YO_low <- 0.0764
strep_coef_YO_high <- 0.1015
  
  
lact_coef_Y <- 0.0089
lact_coef_Y_low <- 0.00635
lact_coef_Y_high <- 0.01145


lact_coef_YO <- 0.00551
lact_coef_YO_low <- 0.00302
lact_coef_YO_high <- 0.008

coefs <- c(strep_coef_Y, strep_coef_Y_low, strep_coef_Y_high, 
  strep_coef_YO, strep_coef_YO_low, strep_coef_YO_high, 
  lact_coef_Y, lact_coef_Y_low, lact_coef_Y_high, 
  lact_coef_YO, lact_coef_YO_low, lact_coef_YO_high)

names <- c("strep_Y", "strep_Y_low", "strep_Y_high", 
  "strep_YO", "strep_YO_low", "strep_YO_high", 
  "lact_Y", "lact_Y_low", "lact_Y_high", 
  "lact_YO", "lact_YO_low", "lact_YO_high")

df <- data.frame(names, coefs)

df$RA_pct <- round((sin(df$coefs))^2 *100,3)
df
#            names   coefs RA_pct
# 1        strep_Y 0.09840  0.965
# 2    strep_Y_low 0.08420  0.707
# 3   strep_Y_high 0.11250  1.260
# 4       strep_YO 0.08890  0.788
# 5   strep_YO_low 0.07640  0.583
# 6  strep_YO_high 0.10150  1.027
# 7         lact_Y 0.00890  0.008
# 8     lact_Y_low 0.00635  0.004
# 9    lact_Y_high 0.01145  0.013
# 10       lact_YO 0.00551  0.003
# 11   lact_YO_low 0.00302  0.001
# 12  lact_YO_high 0.00800  0.006

```





## Reproduce final excel files after filtering taxa depending on prevalence

```{r}
#### need to define the following parameters:
prev_treshold <- 0.2


## From models defined above:
# compa = 

# model <- 
# trans <- 
```


```{r}
savingName <- paste(model, "_", compa, sep="") 

path <- paste(path_analysis, "ResultsBetadiv/", sep="")
res_all <- read.xlsx(file=paste(path, savingName, "_results.xlsx", sep="") , sheetName="res_all")
res_anov_all <- read.xlsx(file=paste(path, savingName, "_results.xlsx", sep="") , sheetName="res_anov_all")
PRS_all <- read.xlsx(file=paste(path, savingName, "_results.xlsx", sep="") , sheetName="PRS_all")

PRS_all_sub <- subset(PRS_all, prev>=prev_treshold)
kept_taxa <- PRS_all_sub$feature
res_anov_all_sub <- subset(res_anov_all, feature %in% kept_taxa)
res_all_sub <- subset(res_all, feature %in% kept_taxa)

PRS_all_sub$q.value <- NULL
res_anov_all_sub$q.value <- NULL
res_all_sub$q.value <- NULL
PRS_all_sub$NA. <- NULL
res_anov_all_sub$NA. <- NULL
res_all_sub$NA. <- NULL

PRS_all_sub$q.value <- p.adjust(PRS_all_sub$p.value, method="fdr")
res_anov_all_sub$q.value <- p.adjust(res_anov_all_sub$p.value, method="fdr")
res_all_sub$q.value <- p.adjust(res_all_sub$p.value, method="fdr")


write.xlsx(res_all_sub, file=paste(path, savingName, "_results_prev", prev_treshold, ".xlsx", sep=""), sheet="res_all")
write.xlsx(res_anov_all_sub, file=paste(path, savingName, "_results_prev", prev_treshold, ".xlsx", sep=""), sheet="res_anov_all", append = T)
write.xlsx(PRS_all_sub, file=paste(path, savingName, "_results_prev", prev_treshold, ".xlsx", sep=""), sheet="PRS_all", append = T)
```






